---
title: "Epigenetic Control of Temperature Response in Ostreococcus tauri (ECTOR)"
author: "Christina Arvanitidou"
date: "2023-05-02"
output: pdf_document
---

## Loading data and preprocessing

```{r}
mapped.reads.k4.10c.1 <- 4660652/1E6 
mapped.reads.k4.10c.2 <- 4572532/1E6
mapped.reads.k4.10c.3 <- 4490999/1E6

mapped.reads.k4.14c.1 <- 4609469/1E6 
mapped.reads.k4.14c.2 <- 5034594/1E6
mapped.reads.k4.14c.3 <- 5174666/1E6

mapped.reads.k4.20c.1 <- 6832850/1E6 
mapped.reads.k4.20c.2 <- 5993486/1E6
mapped.reads.k4.20c.3 <- 6701422/1E6

mapped.reads.k4.24c.1 <- 8444712/1E6 
mapped.reads.k4.24c.2 <- 7768105/1E6
mapped.reads.k4.24c.3 <- 1734986/1E6

mapped.reads.k4.26c.1 <- 6059862/1E6 
mapped.reads.k4.26c.2 <- 5263311/1E6
mapped.reads.k4.26c.3 <- 4251439/1E6

read.counts.k4.temp10c.1 <- read.table(file = "chipseq_data/bed_data/h3k4me3/temp10c/h3k4me3_10_chip_1_count_con24cons.bed", header = F, sep = "\t" )
read.counts.k4.temp10c.2 <- read.table(file = "chipseq_data/bed_data/h3k4me3/temp10c/h3k4me3_10_chip_2_count_con24cons.bed", header = F, sep = "\t" )
read.counts.k4.temp10c.3 <- read.table(file = "chipseq_data/bed_data/h3k4me3/temp10c/h3k4me3_10_chip_3_count_con24cons.bed", header = F, sep = "\t" )

read.counts.k4.temp14c.1 <- read.table(file = "chipseq_data/bed_data/h3k4me3/temp14c/h3k4me3_14_chip_1_count_con24cons.bed", header = F, sep = "\t" )
read.counts.k4.temp14c.2 <- read.table(file = "chipseq_data/bed_data/h3k4me3/temp14c/h3k4me3_14_chip_2_count_con24cons.bed", header = F, sep = "\t" )
read.counts.k4.temp14c.3 <- read.table(file = "chipseq_data/bed_data/h3k4me3/temp14c/h3k4me3_14_chip_3_count_con24cons.bed", header = F, sep = "\t" )

read.counts.k4.temp20c.1 <- read.table(file = "chipseq_data/bed_data/h3k4me3/temp20c/h3k4me3_20_chip_1_count_con24cons.bed", header = F, sep = "\t" )
read.counts.k4.temp20c.2 <- read.table(file = "chipseq_data/bed_data/h3k4me3/temp20c/h3k4me3_20_chip_2_count_con24cons.bed", header = F, sep = "\t" )
read.counts.k4.temp20c.3 <- read.table(file = "chipseq_data/bed_data/h3k4me3/temp20c/h3k4me3_20_chip_3_count_con24cons.bed", header = F, sep = "\t" )

read.counts.k4.temp24c.1 <- read.table(file = "chipseq_data/bed_data/h3k4me3/temp24c/h3k4me3_24_chip_1n_counts.bed", header = F, sep = "\t" )
read.counts.k4.temp24c.2 <- read.table(file = "chipseq_data/bed_data/h3k4me3/temp24c/h3k4me3_24_chip_2n_counts.bed", header = F, sep = "\t" )
read.counts.k4.temp24c.3 <- read.table(file = "chipseq_data/bed_data/h3k4me3/temp24c/h3k4me3_24_chip_3_count.bed", header = F, sep = "\t" )

read.counts.k4.temp26c.1 <- read.table(file = "chipseq_data/bed_data/h3k4me3/temp26c/h3k4me3_26_chip_1_count_con24cons.bed", header = F, sep = "\t" )
read.counts.k4.temp26c.2 <- read.table(file = "chipseq_data/bed_data/h3k4me3/temp26c/h3k4me3_26_chip_2_count_con24cons.bed", header = F, sep = "\t" )
read.counts.k4.temp26c.3 <- read.table(file = "chipseq_data/bed_data/h3k4me3/temp26c/h3k4me3_26_chip_3_count_con24cons.bed", header = F, sep = "\t" )

```


```{r}
peak.cpm <- matrix(data = NA, nrow = nrow(read.counts.k4.temp10c.1), ncol = 18 )
colnames(peak.cpm) <-c("chr","start","end","k4_temp10_1","k4_temp10_2","k4_temp10_3", "k4_temp14_1","k4_temp14_2","k4_temp14_3","k4_temp20_1","k4_temp20_2","k4_temp20_3","k4_temp24_1","k4_temp24_2","k4_temp24_3","k4_temp26_1","k4_temp26_2","k4_temp26_3")

peak.cpm[,1]<- read.counts.k4.temp10c.1[,1]
peak.cpm[,2]<- read.counts.k4.temp10c.1[,2]
peak.cpm[,3]<- read.counts.k4.temp10c.1[,3]

peak.cpm[,4]<- read.counts.k4.temp10c.1[,4]/mapped.reads.k4.10c.1
peak.cpm[,5]<- read.counts.k4.temp10c.2[,4]/mapped.reads.k4.10c.2
peak.cpm[,6]<- read.counts.k4.temp10c.3[,4]/mapped.reads.k4.10c.3

peak.cpm[,7]<- read.counts.k4.temp14c.1[,4]/mapped.reads.k4.14c.1
peak.cpm[,8]<- read.counts.k4.temp14c.2[,4]/mapped.reads.k4.14c.2
peak.cpm[,9]<- read.counts.k4.temp14c.3[,4]/mapped.reads.k4.14c.3

peak.cpm[,10]<- read.counts.k4.temp20c.1[,4]/mapped.reads.k4.20c.1
peak.cpm[,11]<- read.counts.k4.temp20c.2[,4]/mapped.reads.k4.20c.2
peak.cpm[,12]<- read.counts.k4.temp20c.3[,4]/mapped.reads.k4.20c.3

peak.cpm[,13]<- read.counts.k4.temp24c.1[,4]/mapped.reads.k4.24c.1
peak.cpm[,14]<- read.counts.k4.temp24c.2[,4]/mapped.reads.k4.24c.2
peak.cpm[,15]<- read.counts.k4.temp24c.3[,4]/mapped.reads.k4.24c.3


peak.cpm[,16]<- read.counts.k4.temp26c.1[,4]/mapped.reads.k4.26c.1
peak.cpm[,17]<- read.counts.k4.temp26c.2[,4]/mapped.reads.k4.26c.2
peak.cpm[,18]<- read.counts.k4.temp26c.3[,4]/mapped.reads.k4.26c.3

head(peak.cpm)

peak.cpm <- as.data.frame(peak.cpm)

```

#Temperature 10ºC scatterplots
```{r temp10c replicates scatterplots}

plot(x = log2(peak.cpm$k4_temp10_1), y= log2(peak.cpm$k4_temp10_2), pch = 20)
lines(x = c(0,20), y = c(0,20), type = "l", col="blue")

plot(x = log2(peak.cpm$k4_temp10_1), y= log2(peak.cpm$k4_temp10_3), pch = 20)
lines(x = c(0,20), y = c(0,20), type = "l", col="blue")

plot(x = log2(peak.cpm$k4_temp10_2), y= log2(peak.cpm$k4_temp10_3), pch = 20)
lines(x = c(0,20), y = c(0,20), type = "l", col="blue")


boxplot(log2(peak.cpm$k4_temp10_1), log2(peak.cpm$k4_temp10_2), log2(peak.cpm$k4_temp10_3), col= rainbow(3))

```

#Temperature 14ºC scatterplots
```{r temp14c replicates scatterplots}

plot(x = log2(peak.cpm$k4_temp14_1), y= log2(peak.cpm$k4_temp14_2), pch = 20)
lines(x = c(0,20), y = c(0,20), type = "l", col="blue")

plot(x = log2(peak.cpm$k4_temp14_1), y= log2(peak.cpm$k4_temp14_3), pch = 20)
lines(x = c(0,20), y = c(0,20), type = "l", col="blue")

plot(x = log2(peak.cpm$k4_temp14_2), y= log2(peak.cpm$k4_temp14_3), pch = 20)
lines(x = c(0,20), y = c(0,20), type = "l", col="blue")

boxplot(log2(peak.cpm$k4_temp14_1), log2(peak.cpm$k4_temp14_2), log2(peak.cpm$k4_temp14_3), col= rainbow(3))

```


#Temperature 20ºC scatterplots
```{r temp20c replicates scatterplots}

plot(x = log2(peak.cpm$k4_temp20_1), y= log2(peak.cpm$k4_temp20_2), pch = 20)
lines(x = c(0,20), y = c(0,20), type = "l", col="blue")

plot(x = log2(peak.cpm$k4_temp20_1), y= log2(peak.cpm$k4_temp20_3), pch = 20)
lines(x = c(0,20), y = c(0,20), type = "l", col="blue")

plot(x = log2(peak.cpm$k4_temp20_2), y= log2(peak.cpm$k4_temp20_3), pch = 20)
lines(x = c(0,20), y = c(0,20), type = "l", col="blue")

boxplot(log2(peak.cpm$k4_temp20_1), log2(peak.cpm$k4_temp20_2), log2(peak.cpm$k4_temp20_3), col= rainbow(3))

```


#Temperature 24ºC scatterplots
```{r temp24c replicates scatterplots}

plot(x = log2(peak.cpm$k4_temp24_1), y= log2(peak.cpm$k4_temp24_2), pch = 20)
lines(x = c(0,20), y = c(0,20), type = "l", col="blue")

plot(x = log2(peak.cpm$k4_temp24_1), y= log2(peak.cpm$k4_temp24_3), pch = 20)
lines(x = c(0,20), y = c(0,20), type = "l", col="blue")

plot(x = log2(peak.cpm$k4_temp24_2), y= log2(peak.cpm$k4_temp24_3), pch = 20)
lines(x = c(0,20), y = c(0,20), type = "l", col="blue")

boxplot(log2(peak.cpm$k4_temp24_1), log2(peak.cpm$k4_temp24_2), log2(peak.cpm$k4_temp24_3), col= rainbow(3))

```


#Temperature 26ºC scatterplots
```{r temp26c replicates scatterplots}

plot(x = log2(peak.cpm$k4_temp26_1), y= log2(peak.cpm$k4_temp26_2), pch = 20)
lines(x = c(0,20), y = c(0,20), type = "l", col="blue")

plot(x = log2(peak.cpm$k4_temp26_1), y= log2(peak.cpm$k4_temp26_3), pch = 20)
lines(x = c(0,20), y = c(0,20), type = "l", col="blue")

plot(x = log2(peak.cpm$k4_temp26_2), y= log2(peak.cpm$k4_temp26_3), pch = 20)
lines(x = c(0,20), y = c(0,20), type = "l", col="blue")

boxplot(log2(peak.cpm$k4_temp26_1), log2(peak.cpm$k4_temp26_2), log2(peak.cpm$k4_temp26_3), col= rainbow(3))

```

```{r temp10c-temp14c scatterplots}

plot(x = log2(peak.cpm$k4_temp10_1), y= log2(peak.cpm$k4_temp14_1), pch = 20, xlab="Temperature 10ºC",ylab="Temperature 14C",main="14ºC vs 10ºC")
lines(x = c(0,20), y = c(0,20), type = "l", col="blue")

```

```{r temp10c-temp20c scatterplots}

plot(x = log2(peak.cpm$k4_temp10_1), y= log2(peak.cpm$k4_temp20_1), pch = 20, xlab="Temperature 10ºC",ylab="Temperature 20C",main="20ºC vs 10ºC")
lines(x = c(0,20), y = c(0,20), type = "l", col="blue")

```

```{r temp10c-temp24c scatterplots}

plot(x = log2(peak.cpm$k4_temp10_1), y= log2(peak.cpm$k4_temp24_1), pch = 20, xlab="Temperature 10ºC",ylab="Temperature 24C",main="24ºC vs 10ºC")
lines(x = c(0,20), y = c(0,20), type = "l", col="blue")

```


```{r temp10c-temp26c scatterplots}
plot(x = log2(peak.cpm$k4_temp10_1), y= log2(peak.cpm$k4_temp26_1), pch = 20,xlab="Temperature 10ºC",ylab="Temperature 26C",main="26ºC vs 10ºC")
lines(x = c(0,20), y = c(0,20), type = "l", col="blue")



plot(x = log2((peak.cpm$k4_temp10_1 + peak.cpm$k4_temp10_2 +peak.cpm$k4_temp10_3)/3), y= log2((peak.cpm$k4_temp26_1 + peak.cpm$k4_temp26_2 + peak.cpm$k4_temp26_3)/3), pch = 20,xlab="Temperature 10ºC (CPM)",ylab="Temperature 26C (CPM)",main="26ºC vs 10ºC")
lines(x = c(0,20), y = c(0,20), type = "l", col="blue")

boxplot(log2((peak.cpm$k4_temp10_1 + peak.cpm$k4_temp10_2 +peak.cpm$k4_temp10_3)/3),
       log2((peak.cpm$k4_temp26_1 + peak.cpm$k4_temp26_2 + peak.cpm$k4_temp26_3)/3))
```

```{r boxplot global cpm}
boxplot((peak.cpm$k4_temp10_1 + peak.cpm$k4_temp10_2 +peak.cpm$k4_temp10_3)/3,
        (peak.cpm$k4_temp14_1 + peak.cpm$k4_temp14_2 +peak.cpm$k4_temp14_3)/3,
        (peak.cpm$k4_temp20_1 + peak.cpm$k4_temp20_2 +peak.cpm$k4_temp20_3)/3,
        (peak.cpm$k4_temp26_1 + peak.cpm$k4_temp26_2+peak.cpm$k4_temp26_3)/3,
        outline=F,col=c("#00FFFF","#6699FF","#CC33FF","#FF0000"),ylab="CPM",names=c("10ºC","14ºC","20ºC","26ºC"),las=2,cex.lab=2)

wilcox.test(x = (peak.cpm$k4_temp10_1 + peak.cpm$k4_temp10_2 +peak.cpm$k4_temp10_3)/3, y = (peak.cpm$k4_temp14_1 + peak.cpm$k4_temp14_2 +peak.cpm$k4_temp14_3)/3,alternative = "less")

wilcox.test(x = (peak.cpm$k4_temp14_1 + peak.cpm$k4_temp14_2 +peak.cpm$k4_temp14_3)/3, y = (peak.cpm$k4_temp20_1 + peak.cpm$k4_temp20_2 +peak.cpm$k4_temp20_3)/3,alternative = "less")

wilcox.test(x = (peak.cpm$k4_temp10_1 + peak.cpm$k4_temp10_2 +peak.cpm$k4_temp10_3)/3, y = (peak.cpm$k4_temp20_1 + peak.cpm$k4_temp20_2 +peak.cpm$k4_temp20_3)/3,alternative = "less")

wilcox.test(x = (peak.cpm$k4_temp20_1 + peak.cpm$k4_temp20_2 +peak.cpm$k4_temp20_3)/3, y = (peak.cpm$k4_temp26_1 + peak.cpm$k4_temp26_2 +peak.cpm$k4_temp26_3)/3,alternative = "less")




boxplot((peak.cpm$k4_temp10_1 + peak.cpm$k4_temp10_2 +peak.cpm$k4_temp10_3)/3,
        (peak.cpm$k4_temp14_1 + peak.cpm$k4_temp14_2 +peak.cpm$k4_temp14_3)/3,
        (peak.cpm$k4_temp20_1 + peak.cpm$k4_temp20_2 +peak.cpm$k4_temp20_3)/3,
        (peak.cpm$k4_temp24_1 + peak.cpm$k4_temp24_2 +peak.cpm$k4_temp24_3)/3,
        (peak.cpm$k4_temp26_1 + peak.cpm$k4_temp26_2 +peak.cpm$k4_temp26_3)/3,
        outline=F,col=c("#00FFFF","#6699FF","#CC33FF","#FF0000","blue"),ylab="CPM",names=c("10ºC","14ºC","20ºC","24ºC","26ºC"),las=2,cex.lab=2)

```

```{r}

library(NormalyzerDE)

dim(peak.cpm)

write.table(x = peak.cpm[4:18],file = "peak_cpm_k4.tsv" ,
            quote = F,row.names = F,
            sep = "\t")

design.k4 <- data.frame(sample=colnames(peak.cpm[4:18]),
                     group=c(rep("temp10",3),rep("temp14",3), rep("temp20",3), rep("temp24",3), rep("temp26",3)))



write.table(x = design.k4,file = "normalyzer_design_k4_global.tsv",quote = F,row.names = F,
            sep = "\t")


normalyzer(jobName = "normalyzer_design_k4_global",designPath = "normalyzer_design_k4_global.tsv",
           dataPath = "peak_cpm_k4.tsv",outputDir = ".", omitLowAbundSamples = T)


normalized.peak.cpm <- read.table(file="normalyzer_design_k4_global/Quantile-normalized.txt", header=T)
head(normalized.peak.cpm)


boxplot(normalized.peak.cpm, outline=F, cex.lab=1.5)


```

```{r}

head(normalized.peak.cpm)

peak.cpm[,1]<- read.counts.k4.temp10c.1[,1]
peak.cpm[,2]<- read.counts.k4.temp10c.1[,2]
peak.cpm[,3]<- read.counts.k4.temp10c.1[,3]


normalized.peak.cpm <-  cbind(peak.cpm[,1], peak.cpm[,2], peak.cpm[,3], normalized.peak.cpm)
colnames(normalized.peak.cpm)[1:3] <- c("chr", "start", "end")
head(normalized.peak.cpm)



boxplot((normalized.peak.cpm$k4_temp10_1 + normalized.peak.cpm$k4_temp10_2 +normalized.peak.cpm$k4_temp10_3)/3,
        (normalized.peak.cpm$k4_temp14_1 + normalized.peak.cpm$k4_temp14_2 +normalized.peak.cpm$k4_temp14_3)/3,
       (normalized.peak.cpm$k4_temp20_1 + normalized.peak.cpm$k4_temp20_2 +normalized.peak.cpm$k4_temp20_3)/3,
       (normalized.peak.cpm$k4_temp24_1 + normalized.peak.cpm$k4_temp24_2 +normalized.peak.cpm$k4_temp24_3)/3,
       (normalized.peak.cpm$k4_temp26_1 + normalized.peak.cpm$k4_temp26_2 +normalized.peak.cpm$k4_temp26_3)/3,
      outline=F,col=c("#00FFFF","#6699FF","#CC33FF","#FF0000","blue"),ylab="CPM",names=c("10ºC","14ºC","20ºC","24ºC","26ºC"),las=2,cex.lab=2)





```
```{r}
normalized.peak.names <- paste(paste(normalized.peak.cpm$chr, normalized.peak.cpm$start, sep = "_"), normalized.peak.cpm$end, sep = "_")
mat.normalized.peak.cpm <- as.matrix(normalized.peak.cpm[,4:18])
rownames(mat.normalized.peak.cpm) <- normalized.peak.names
head(mat.normalized.peak.cpm)
```

#Creating matrix 
```{r}
#peak.names <- paste(paste(peak.cpm$chr, peak.cpm$start, sep = "_"), peak.cpm$end, sep = "_")
#log2.peak.cpm <- matrix(data = NA, nrow = nrow(read.counts.k4.temp10c.1), ncol = 15)
#colnames(log2.peak.cpm) <- c("k4_temp10_1","k4_temp10_2","k4_temp10_3", "k4_temp14_1","k4_temp14_2","k4_temp14_3","k4_temp20_1","k4_temp20_2","k4_temp20_3","k4_temp24_1","k4_temp24_2","k4_temp24_3","k4_temp26_1","k4_temp26_2","k4_temp26_3")
#rownames(log2.peak.cpm) <- peak.names 

#log2.peak.cpm[,"k4_temp10_1"] <- log2(peak.cpm$k4_temp10_1)
#log2.peak.cpm[,"k4_temp10_2"] <- log2(peak.cpm$k4_temp10_2)
#log2.peak.cpm[,"k4_temp10_3"] <- log2(peak.cpm$k4_temp10_3)

#log2.peak.cpm[,"k4_temp14_1"] <- log2(peak.cpm$k4_temp14_1)
#log2.peak.cpm[,"k4_temp14_2"] <- log2(peak.cpm$k4_temp14_2)
#log2.peak.cpm[,"k4_temp14_3"] <- log2(peak.cpm$k4_temp14_3)

#log2.peak.cpm[,"k4_temp20_1"] <- log2(peak.cpm$k4_temp20_1)
#log2.peak.cpm[,"k4_temp20_2"] <- log2(peak.cpm$k4_temp20_2)
#log2.peak.cpm[,"k4_temp20_3"] <- log2(peak.cpm$k4_temp20_3)

#log2.peak.cpm[,"k4_temp24_1"] <- log2(peak.cpm$k4_temp24_1)
#log2.peak.cpm[,"k4_temp24_2"] <- log2(peak.cpm$k4_temp24_2)
#log2.peak.cpm[,"k4_temp24_3"] <- log2(peak.cpm$k4_temp24_3)

#log2.peak.cpm[,"k4_temp26_1"] <- log2(peak.cpm$k4_temp26_1)
#log2.peak.cpm[,"k4_temp26_2"] <- log2(peak.cpm$k4_temp26_2)
#log2.peak.cpm[,"k4_temp26_3"] <- log2(peak.cpm$k4_temp26_3)


#head(log2.peak.cpm)

#boxplot(log2.peak.cpm, las=2)

```


```{r}
library(limma)

experimental.design <- model.matrix(~ -1+factor(c(1,1,1,2,2,2,3,3,3,4,4,4,5,5,5)))
colnames(experimental.design) <- c("k4_temp10","k4_temp14","k4_temp20","k4_temp24","k4_temp26")

experimental.design

linear.fit <- lmFit(mat.normalized.peak.cpm, experimental.design)

contrast.matrix <- makeContrasts(k4_temp14-k4_temp10,k4_temp20-k4_temp10,k4_temp24-k4_temp10,k4_temp26-k4_temp10,levels=c("k4_temp10","k4_temp14","k4_temp20","k4_temp24","k4_temp26"))

contrast.linear.fit <- contrasts.fit(linear.fit, contrast.matrix)
contrast.results <- eBayes(contrast.linear.fit)

```

```{r temp14c-temp10c contrast}
temp14c.vs.temp10c <- topTable(contrast.results, number=nrow(mat.normalized.peak.cpm),coef=1,sort.by="logFC")
head(temp14c.vs.temp10c)
peak.ids.14c.10c <- rownames(temp14c.vs.temp10c) 
fold.change.14c.10c <- temp14c.vs.temp10c$logFC
q.value.14c.10c <- temp14c.vs.temp10c$adj.P.Val

activated.peaks.temp14c.vs.temp10c <- peak.ids.14c.10c[fold.change.14c.10c > 1 & q.value.14c.10c < 0.05]
repressed.peaks.temp14c.vs.temp10c <- peak.ids.14c.10c[fold.change.14c.10c < - 1 & q.value.14c.10c < 0.05]

length(activated.peaks.temp14c.vs.temp10c)
length(repressed.peaks.temp14c.vs.temp10c)

```


```{r temp20c-temp10c contrast}
temp20c.vs.temp10c <- topTable(contrast.results, number=nrow(mat.normalized.peak.cpm),coef=2,sort.by="logFC")
head(temp20c.vs.temp10c)
peak.ids.20c.10c <- rownames(temp20c.vs.temp10c) 
fold.change.20c.10c <- temp20c.vs.temp10c$logFC
q.value.20c.10c <- temp20c.vs.temp10c$adj.P.Val

activated.peaks.temp20c.vs.temp10c <- peak.ids.20c.10c[fold.change.20c.10c > 1 & q.value.20c.10c < 0.05]
repressed.peaks.temp20c.vs.temp10c <- peak.ids.20c.10c[fold.change.20c.10c < - 1 & q.value.20c.10c < 0.05]

length(activated.peaks.temp20c.vs.temp10c)
length(repressed.peaks.temp20c.vs.temp10c)


```

```{r temp24c-temp10c contrast}
temp24c.vs.temp10c <- topTable(contrast.results, number=nrow(mat.normalized.peak.cpm),coef=3,sort.by="logFC")
head(temp24c.vs.temp10c)
peak.ids.24c.10c <- rownames(temp24c.vs.temp10c) 
fold.change.24c.10c <- temp24c.vs.temp10c$logFC
q.value.24c.10c <- temp24c.vs.temp10c$adj.P.Val

activated.peaks.temp24c.vs.temp10c <- peak.ids.24c.10c[fold.change.24c.10c > 1 & q.value.24c.10c < 0.05]
repressed.peaks.temp24c.vs.temp10c <- peak.ids.24c.10c[fold.change.24c.10c < - 1 & q.value.24c.10c < 0.05]

length(activated.peaks.temp24c.vs.temp10c)
length(repressed.peaks.temp24c.vs.temp10c)

```


```{r temp26c-temp10c contrast}
temp26c.vs.temp10c <- topTable(contrast.results, number=nrow(mat.normalized.peak.cpm),coef=4,sort.by="logFC")
head(temp26c.vs.temp10c)
peak.ids.26c.10c <- rownames(temp26c.vs.temp10c) 
fold.change.26c.10c <- temp26c.vs.temp10c$logFC
q.value.26c.10c <- temp26c.vs.temp10c$adj.P.Val


activated.peaks.temp26c.vs.temp10c <- peak.ids.26c.10c[fold.change.26c.10c > 1 & q.value.26c.10c < 0.05]
repressed.peaks.temp26c.vs.temp10c <- peak.ids.26c.10c[fold.change.26c.10c < - 1 & q.value.26c.10c < 0.05]

length(activated.peaks.temp26c.vs.temp10c)
length(repressed.peaks.temp26c.vs.temp10c)


length(peak.names)
length(activated.peaks.temp26c.vs.temp10c)/length(peak.names)

strsplit(x = activated.peaks.temp26c.vs.temp10c, split = "_")
matrix.activated.peaks.temp26c.vs.temp10c <- as.matrix(t(as.data.frame(strsplit(x = activated.peaks.temp26c.vs.temp10c, split = "_"))))
head(matrix.activated.peaks.temp26c.vs.temp10c)
rownames(matrix.activated.peaks.temp26c.vs.temp10c) <- NULL
head(matrix.activated.peaks.temp26c.vs.temp10c)

activated.peaks.temp26c.vs.temp10c[1:3]
head(matrix.activated.peaks.temp26c.vs.temp10c, n=3)

write.table(matrix.activated.peaks.temp26c.vs.temp10c, file = "chipseq_results/h3k4me3/limma.k4.activated.peaks.temp26.vs.temp10.bed", sep = "\t", quote = FALSE, row.names = FALSE, col.names = FALSE)

library(ChIPseeker)
library(TxDb.Otauri.JGI)
txdb <- TxDb.Otauri.JGI

temp26.vs.temp10.peaks <- readPeakFile(peakfile = "chipseq_results/h3k4me3/limma.k4.activated.peaks.temp26.vs.temp10.bed",header=FALSE)
promoter <- getPromoters(TxDb=txdb, 
                         upstream=1000, 
                         downstream=1000)
temp26.vs.temp10.peakAnno <- annotatePeak(peak = temp26.vs.temp10.peaks, 
                             tssRegion=c(-1000, 1000),
                             TxDb=txdb)

plotAnnoPie(temp26.vs.temp10.peakAnno)
plotAnnoBar(temp26.vs.temp10.peakAnno)
plotDistToTSS(temp26.vs.temp10.peakAnno,
              title="Distribution of genomic loci relative to TSS",
              ylab = "Genomic Loci (%) (5' -> 3')")

temp26.vs.temp10.annotation <- as.data.frame(temp26.vs.temp10.peakAnno)
head(temp26.vs.temp10.annotation)
length(temp26.vs.temp10.annotation$geneId)

write(x = temp26.vs.temp10.annotation$geneId,file = "chipseq_results/h3k4me3/limma.k4.temp26.vs.temp10.activated.peaks.genes.txt")



ostta2atha.deseq.26.vs.10.activated.genes.peaks<- read.table(file = "chipseq_results/h3k4me3/ostta2atha_deseq_k4_26_vs_10_activated_peak_genes.txt", header = F)
ostta2atha.deseq.26.vs.10.activated.genes.peaks <-ostta2atha.deseq.26.vs.10.activated.genes.peaks[,2]
write(x = ostta2atha.deseq.26.vs.10.activated.genes.peaks,file = "chipseq_results/h3k4me3/ostta2atha.final.deseq.26.vs.10.activated.genes.peaks.txt")


```



```{r}
#xpos <- boxplot(rowMeans(2^log2.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp10",1:3,sep="_")]),
#rowMeans(2^log2.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp14",1:3,sep="_")]),
#rowMeans(2^log2.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp20",1:3,sep="_")]),
#rowMeans(2^log2.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp26",1:3,sep="_")]),outline=F,col=c("#00FFFF","#6699FF", "#CC33FF","#FF0000"),ylab="CPM",names=c("10ºC","14ºC","20ºC","26ºC"),las=2,cex.lab=1.2, main= "H3K4me3 Differentially Marked Genes" )

#wilcox.test(x = rowMeans(2^log2.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp10",1:3,sep="_")]), y = rowMeans(2^log2.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp14",1:3,sep="_")]),alternative = "less")

#wilcox.test(x = rowMeans(2^log2.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp14",1:3,sep="_")]), y = rowMeans(2^log2.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp20",1:3,sep="_")]),alternative = "less")

#wilcox.test(x = rowMeans(2^log2.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp10",1:3,sep="_")]), y = rowMeans(2^log2.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp20",1:3,sep="_")]),alternative = "less")

#wilcox.test(x = rowMeans(2^log2.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp10",1:3,sep="_")]), y = rowMeans(2^log2.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp26",1:3,sep="_")]),alternative = "less")

#wilcox.test(x = rowMeans(2^log2.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp20",1:3,sep="_")]), y = rowMeans(2^log2.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp26",1:3,sep="_")]),alternative = "less")

#wilcox.test(x = rowMeans(2^log2.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp10",1:3,sep="_")]), y = rowMeans(2^log2.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp24",1:3,sep="_")]),alternative = "less")

#wilcox.test(x = rowMeans(2^log2.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp24",1:3,sep="_")]), y = rowMeans(2^log2.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp20",1:3,sep="_")]),alternative = "less")

#wilcox.test(x = rowMeans(2^log2.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp24",1:3,sep="_")]), y = rowMeans(2^log2.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp26",1:3,sep="_")]),alternative = "less")


#boxplot(rowMeans(2^log2.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp10",1:3,sep="_")]),
#rowMeans(2^log2.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp14",1:3,sep="_")]),
#rowMeans(2^log2.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp20",1:3,sep="_")]),
#rowMeans(2^log2.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp24",1:3,sep="_")]),
#rowMeans(2^log2.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp26",1:3,sep="_")]),outline=F,col=c("#00FFFF","#6699FF", "#CC33FF","#FF0000", "grey"),ylab="CPM",names=c("10ºC","14ºC","20ºC","24ºC","26ºC"),las=2,cex.lab=1)



target.genes <- temp26.vs.temp10.annotation$geneId

boxplot(rowMeans(expression.matrix[target.genes,paste("T10",1:3,sep="_")]),
rowMeans(expression.matrix[target.genes,paste("T14",1:3,sep="_")]),
rowMeans(expression.matrix[target.genes,paste("T20",1:3,sep="_")]),
rowMeans(expression.matrix[target.genes,paste("T26",1:3,sep="_")]),outline=F,col=c("#00FFFF","#6699FF", "#CC33FF","#FF0000"),ylab="FPKM",names=c("10ºC","14ºC","20ºC","26ºC"),las=2,cex.lab=1.2, main = "H3K4me3 Differentially Marked Genes" )

head(expression.matrix)
```


```{r}
boxplot(rowMeans(2^mat.normalized.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp10",1:3,sep="_")]),
rowMeans(2^mat.normalized.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp14",1:3,sep="_")]),
rowMeans(2^mat.normalized.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp20",1:3,sep="_")]),
rowMeans(2^mat.normalized.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp24",1:3,sep="_")]),
rowMeans(2^mat.normalized.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp26",1:3,sep="_")]),outline=F,col=c("#00FFFF","#6699FF", "#CC33FF","#FF0000", "grey"),ylab="CPM",names=c("10ºC","14ºC","20ºC","24ºC","26ºC"),las=2,cex.lab=1)


wilcox.test(x = rowMeans(2^mat.normalized.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp10",1:3,sep="_")]), y = rowMeans(2^log2.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp14",1:3,sep="_")]),alternative = "less")

wilcox.test(x = rowMeans(2^mat.normalized.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp14",1:3,sep="_")]), y = rowMeans(2^log2.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp20",1:3,sep="_")]),alternative = "less")

wilcox.test(x = rowMeans(2^mat.normalized.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp10",1:3,sep="_")]), y = rowMeans(2^log2.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp20",1:3,sep="_")]),alternative = "less")

wilcox.test(x = rowMeans(2^mat.normalized.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp10",1:3,sep="_")]), y = rowMeans(2^log2.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp26",1:3,sep="_")]),alternative = "less")

wilcox.test(x = rowMeans(2^mat.normalized.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp20",1:3,sep="_")]), y = rowMeans(2^log2.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp26",1:3,sep="_")]),alternative = "less")

wilcox.test(x = rowMeans(2^mat.normalized.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp10",1:3,sep="_")]), y = rowMeans(2^log2.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp24",1:3,sep="_")]),alternative = "less")

wilcox.test(x = rowMeans(2^mat.normalized.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp20",1:3,sep="_")]), y = rowMeans(2^log2.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp24",1:3,sep="_")]),alternative = "less")

wilcox.test(x = rowMeans(2^mat.normalized.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp24",1:3,sep="_")]), y = rowMeans(2^log2.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp26",1:3,sep="_")]),alternative = "less")

wilcox.test(x = rowMeans(2^mat.normalized.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp20",1:3,sep="_")]), y = rowMeans(2^log2.peak.cpm[activated.peaks.temp26c.vs.temp10c,paste("k4_temp26",1:3,sep="_")]),alternative = "less")

```






```{r limma PCA analysis}
library(tidyverse)
library(FactoMineR)
library("factoextra")
pca.peak.counts <- data.frame(colnames(log2.peak.cpm),t(log2.peak.cpm))
colnames(pca.peak.counts)[1] <- "temperature"

res.pca <- PCA(pca.peak.counts, graph = FALSE,scale.unit = TRUE,quali.sup = 1 )
png(filename = "chipseq_results/h3k4me3/k4.pca_analysis_1_limma.png")
fviz_eig(res.pca, addlabels = TRUE, ylim = c(0, 70),main = "")

pheno.data <- read.table(file = "pheno_data_k4.csv", header = T, sep = ",")
pheno.data

png(filename = "chipseq_results/h3k4me3/pca_analysis_2_limma.png")
fviz_pca_ind(res.pca, col.ind = pheno.data$temperature[1:15], 
             pointsize=2, pointshape=21,fill="black",
             repel = TRUE, 
             addEllipses = TRUE,ellipse.type = "confidence",
             legend.title="Conditions",
             title="Principal Components Analysis H3K4me3",
             show_legend=TRUE,show_guide=TRUE) 


```


```{r}
ostta2atha <- read.table(file = "chipseq_results/h3k4me3/raw_ostta2atha.tsv", header = F)
ostta.limma.26.10.genes <-  read.table(file = "chipseq_results/h3k4me3/limma.k4.temp26.vs.temp10.activated.peaks.genes.txt", header = F)


#ostta2atha.limma.26.10.genes <- ostta2atha$V2[ostta.limma.26.10.genes$V1 == ostta2atha$V1]
ostta2atha.limma.26.10.genes <- ostta2atha[ostta2atha$V1 %in% ostta.limma.26.10.genes$V1, ]
ostta2atha.limma.26.10.genes <- ostta2atha.limma.26.10.genes[,1:2]
length(ostta2atha.limma.26.10.genes)
rownames(ostta2atha.limma.26.10.genes) <- NULL
head(ostta2atha.limma.26.10.genes)
dim(ostta2atha.limma.26.10.genes)
write.table(x = ostta2atha.limma.26.10.genes, file = "chipseq_results/h3k4me3/ostta2atha.limma.26.10.genes.txt", sep = "\t")


```
stress response and heat shock proteins
AT1G64550.1 ostta06g03620
AT4G17090.1 ostta02g04570
AT4G17040.1 ostta02g00380
AT1G27980.1 ostta02g02240
AT5G56000.1 ostta10g01280 hsp ??? 
AT3G20670.1 ostta16g01060
AT1G62740.1 ostta09g01040 
AT4G37930.1 ostta12g01860
AT3G47550   ostta10g00190

general stress
ostta01g03420 AT1G15690.1 

ostta10g00190	AT3G47550.1

ostta10g00940	AT5G50200.3

ostta12g01000	AT5G27600.1

ostta12g01860	AT4G37930.1

ostta13g00885	AT4G04330.1

ostta17g00880	AT4G34220.1

ostta18g01560	AT1G02890.2

ostta02g00220	AT4G30250.1

ostta02g00380	AT4G17040.1

ostta02g00690	AT5G46280.1

ostta02g02710	AT2G18950.1

ostta02g02940	AT5G51160.1 ??

ostta02g04570	AT4G17090.1

ostta03g00240	AT1G63830.5 ??

ostta03g01360	AT3G10800.1

ostta04g00130	AT1G55480.1

ostta04g03510	AT2G43710.2

ostta06g03620	AT1G64550.1
	
ostta07g01550	AT4G23730.1 ??

ostta07g03640	AT3G54440.1

ostta08g02280	AT3G03540.1

ostta09g01010	AT1G72760.1









fatty acid metabolism
AT3G03540.1 ostta08g02280
AT1G62740.1 ostta09g01040
AT2G43710.1 ostta04g03510?




```{r DESeq2}
library(DESeq2)

peak.counts <- matrix(data = NA, nrow = nrow(read.counts.k4.temp10c.1), ncol = 15)
colnames(peak.counts) <- c("k4_temp10_1","k4_temp10_2","k4_temp10_3", "k4_temp14_1","k4_temp14_2","k4_temp14_3","k4_temp20_1","k4_temp20_2","k4_temp20_3","k4_temp24_1","k4_temp24_2","k4_temp24_3","k4_temp26_1","k4_temp26_2","k4_temp26_3")
rownames(peak.counts) <- peak.names

peak.counts[,"k4_temp10_1"] <- read.counts.k4.temp10c.1$V4
peak.counts[,"k4_temp10_2"] <- read.counts.k4.temp10c.2$V4
peak.counts[,"k4_temp10_3"] <- read.counts.k4.temp10c.3$V4
peak.counts[,"k4_temp14_1"] <- read.counts.k4.temp14c.1$V4
peak.counts[,"k4_temp14_2"] <- read.counts.k4.temp14c.2$V4
peak.counts[,"k4_temp14_3"] <- read.counts.k4.temp14c.3$V4
peak.counts[,"k4_temp20_1"] <- read.counts.k4.temp20c.1$V4
peak.counts[,"k4_temp20_2"] <- read.counts.k4.temp20c.2$V4
peak.counts[,"k4_temp20_3"] <- read.counts.k4.temp20c.3$V4
peak.counts[,"k4_temp24_1"] <- read.counts.k4.temp24c.1$V4
peak.counts[,"k4_temp24_2"] <- read.counts.k4.temp24c.2$V4
peak.counts[,"k4_temp24_3"] <- read.counts.k4.temp24c.3$V4
peak.counts[,"k4_temp26_1"] <- read.counts.k4.temp26c.1$V4
peak.counts[,"k4_temp26_2"] <- read.counts.k4.temp26c.2$V4
peak.counts[,"k4_temp26_3"] <- read.counts.k4.temp26c.3$V4

head(peak.counts)
pheno.data <- read.table(file = "pheno_data_k4.csv", header = T, sep = ",")
pheno.data

dds <- DESeqDataSetFromMatrix(countData=peak.counts, colData=pheno.data, design = ~ temperature)

dds <- DESeq(dds) 
```

```{r deseq2 temp14 vs temp10}
temp14vstemp10 <- results(dds, contrast = c("temperature", "temp14", "temp10"))

log.fold.change <- temp14vstemp10$log2FoldChange
q.value <- temp14vstemp10$padj
names(log.fold.change) <- rownames(temp14vstemp10)
names(q.value) <- rownames(temp14vstemp10)

activated.peaks.14.10.deseq2 <- rownames(temp14vstemp10)[log.fold.change > 1 & q.value < 0.05]
length(activated.peaks.14.10.deseq2)

repressed.genes.14.10.deseq2 <- rownames(temp14vstemp10)[log.fold.change < -  1 & q.value < 0.05]
length(repressed.genes.14.10.deseq2)

```

```{r deseq2 temp20 vs temp10}
temp20vstemp10 <- results(dds, contrast = c("temperature", "temp20", "temp10"))

log.fold.change <- temp20vstemp10$log2FoldChange
q.value <- temp20vstemp10$padj
names(log.fold.change) <- rownames(temp20vstemp10)
names(q.value) <- rownames(temp20vstemp10)

activated.peaks.20.10.deseq2 <- rownames(temp20vstemp10)[log.fold.change > 1 & q.value < 0.05]
length(activated.peaks.20.10.deseq2)
write(x = activated.peaks.20.10.deseq2, file = "chipseq_results/h3k4me3/deseq2.k4.activated.peaks.20.10.txt", sep = "_")

repressed.genes.20.10.deseq2 <- rownames(temp20vstemp10)[log.fold.change < -  1 & q.value < 0.05]
length(repressed.genes.20.10.deseq2)

```

```{r deseq2 temp24 vs temp10}
temp24vstemp10 <- results(dds, contrast = c("temperature", "temp24", "temp10"))

log.fold.change <- temp24vstemp10$log2FoldChange
q.value <- temp24vstemp10$padj
names(log.fold.change) <- rownames(temp24vstemp10)
names(q.value) <- rownames(temp24vstemp10)

activated.peaks.24.10.deseq2 <- rownames(temp24vstemp10)[log.fold.change > 1 & q.value < 0.05]
length(activated.peaks.24.10.deseq2)
write(x = activated.peaks.24.10.deseq2, file = "chipseq_results/h3k4me3/deseq2.k4.activated.peaks.24.10.txt", sep = "_")

repressed.genes.24.10.deseq2 <- rownames(temp24vstemp10)[log.fold.change < -  1 & q.value < 0.05]
length(repressed.genes.24.10.deseq2)

```


```{r deseq2 temp26 vs temp10}
temp26vstemp10 <- results(dds, contrast = c("temperature", "temp26", "temp10"))

log.fold.change <- temp26vstemp10$log2FoldChange
q.value <- temp26vstemp10$padj
names(log.fold.change) <- rownames(temp26vstemp10)
names(q.value) <- rownames(temp26vstemp10)

activated.peaks.26.10.deseq2 <- rownames(temp26vstemp10)[log.fold.change > 1 & q.value < 0.05]
length(activated.peaks.26.10.deseq2)

repressed.genes.26.10.deseq2 <- rownames(temp26vstemp10)[log.fold.change < -  1 & q.value < 0.05]
length(repressed.genes.26.10.deseq2)

strsplit(x = activated.peaks.26.10.deseq2, split = "_")
matrix.activated.peaks.26.10.deseq2 <- as.matrix(t(as.data.frame(strsplit(x = activated.peaks.26.10.deseq2, split = "_"))))
head(matrix.activated.peaks.26.10.deseq2)
rownames(matrix.activated.peaks.26.10.deseq2) <- NULL
head(matrix.activated.peaks.26.10.deseq2)

matrix.activated.peaks.26.10.deseq2[1:3]
head(matrix.activated.peaks.26.10.deseq2, n=3)

write.table(matrix.activated.peaks.26.10.deseq2, file = "chipseq_results/h3k4me3/deseq2.k4.activated.peaks.26.10.bed", sep = "\t", quote = FALSE, row.names = FALSE, col.names = FALSE)


deseq.temp26.vs.temp10.peaks <- readPeakFile(peakfile = "chipseq_results/h3k4me3/deseq2.k4.activated.peaks.26.10.bed", header=FALSE)
promoter <- getPromoters(TxDb=txdb, 
                         upstream=1000, 
                         downstream=1000)
deseq.temp26.vs.temp10.peakAnno <- annotatePeak(peak = deseq.temp26.vs.temp10.peaks, 
                             tssRegion=c(-1000, 1000),
                             TxDb=txdb)

plotAnnoPie(deseq.temp26.vs.temp10.peakAnno)
plotAnnoBar(deseq.temp26.vs.temp10.peakAnno)
plotDistToTSS(deseq.temp26.vs.temp10.peakAnno,
              title="Distribution of genomic loci relative to TSS",
              ylab = "Genomic Loci (%) (5' -> 3')")

deseq.temp26.vs.temp10.annotation <- as.data.frame(deseq.temp26.vs.temp10.peakAnno)
head(deseq.temp26.vs.temp10.annotation)


write(x = deseq.temp26.vs.temp10.annotation$geneId,file = "chipseq_results/h3k4me3/deseq.k4.temp26.vs.temp10.activated.peaks.genes.txt")

```

ostta03g01360	AT3G10800 heat response in atha
ostta13g00885 AT4G04330 chaperone


ostta12g01900	length:700 (mRNA) (n/a)  (Carotene isomerase)
ostta06g02290	length:432 (mRNA) (n/a)  (Cobalamin (vitamin B12) biosynthesis CobW-like, C-terminal)
ostta02g00420	length:485 (mRNA) (n/a)  (Kinetochore protein Ndc80) (n/a)  (Kinetochore protein Ndc80)
ostta07g03640
ostta18g01570	length:375 (mRNA) (n/a)  (Cyclin A/B/D/E)




#####Volcano plot






