---
title: "Epigenetic Control of Temperature Response in Ostreococcus tauri (ECTOR)"
author: "Christina Arvanitidou (carvanitidou@us.es) and Francisco J. Romero-Campero (fran@us.es)"
date: '2024-03-15'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## H3K27me3 genome wide distribution analysis

We load the genome of Ostreococcus tauri and compute the length of its chromosomes. We determine the length of the largest chromosome and for technical reasons we reverse the vector of lengths. This will be used to represent graphically the genome wide distribution of H3K27me3. 

```{r load_genome}
library(seqinr)

ostta.genome <- read.fasta(file = "ostta_genome/ostreococcus_tauri.fa",seqtype = "DNA")
chr.lens <- sapply(X = getSequence(ostta.genome),FUN = length)

number.chrs <- length(getSequence(ostta.genome))
max.chr.len <- max(chr.lens)
chr.lens <- rev(chr.lens)
```

We load the H3K27me3 peaks identified in cultures grown under constant light at 20ºC. We represent each chromosome using a rectangle. Each H3K27me3 peak is represented in its precise genome location using a blue rectangle. 

```{r visualization of H3K27me3 genome wide distribution}
ll.20 <- read.table(file="chipseq_data/bed_data/h3k27me3/h3k27me3_ll_20.bed",
                       header = F)

ll.20 <- ll.20[,1:3]

nrow(ll.20)

#jpeg(filename = "images/h3k27me3_genome_wide_distribution_ll_20.jpg",res = 300,width = 900,height = 900)
#par(mar=c(0,0,0,0))
plot(x=c(0,max.chr.len), y=c(0,4*number.chrs), col = "white", xlab = "", ylab = "", axes=F)  

for(i in 1:number.chrs)
{
 polygon(x = c(0, chr.lens[i], chr.lens[i], 0),   
         y = c(4*i+1, 4*i+1, 4*i-1, 4*i-1),    
         col = "white") 
}

for(i in 1:nrow(ll.20))
{
 current.chr <- ll.20[i,1]
 current.start <- ll.20[i,2]
 current.end <- ll.20[i,3]

 current.line <- (20:1)[current.chr]
 
 polygon(x = c(current.start, current.end, current.end, current.start),   
         y = c(4*current.line+1, 4*current.line+1, 4*current.line-1, 4*current.line-1),    
         col = "blue",border="blue")
}
#dev.off()
```

We identified 480 genomic regions significantly occupied by H3K27me3 in the three different independent biological replicates. We compute the percentage of the whole genome occupied by H3K27me3 and the percentage in each chromosome. A barplot is used to represent the percentage of H3K27me3 in each chromosome. Z-scores and p-values are computed for each chromosome. A gradient from white to blue is used to represent significance levels.  

```{r k27_in_chrs}
h3k27me3.peaks.len <- ll.20$V3 - ll.20$V2
sum(h3k27me3.peaks.len)/sum(chr.lens)
ll.20 <- cbind(ll.20,h3k27me3.peaks.len)
colnames(ll.20) <- c("chr","start","end","length")
head(ll.20)
number.of.h3k27.peaks.chr <- vector(mode = "numeric",length = 20)
peaks.len.chr <- vector(mode = "numeric",length = 20)

for(i in 1:20)
{
 number.of.h3k27.peaks.chr[i] <- nrow(subset(ll.20, chr == i))
 peaks.len.chr[i] <- sum(subset(ll.20, chr == i)[,4]) 
}

percentage.chr.k27 <- peaks.len.chr/rev(chr.lens)

z.scores <- (percentage.chr.k27 - mean(percentage.chr.k27))/sd(percentage.chr.k27)

which(z.scores > 2.5)

p.vals <- pnorm((percentage.chr.k27 - mean(percentage.chr.k27[-19]))/sd(percentage.chr.k27[-19]),lower.tail = F)

p.vals

values.for.colors <- round(-log10(p.vals)*40)
values.for.colors[19] <- 100

white.blue.palette <- colorRampPalette(c('white','blue'))

#jpeg(filename = "images/h3k27me3_chr_barplot_ll_20.jpg",res = 300,width = 1800,height = 900)
barplot(100*percentage.chr.k27,border="black",names.arg = paste0("Chr",1:20),las=2,col=white.blue.palette(100)[values.for.colors]
)
abline(h = 100*mean(percentage.chr.k27[-19]),lty=5)
#dev.off()
```

Approximately 7.2% of the Ostreococcus genome is occupied by H3K27me3. 
H3K27me3 is mostly uniformly distributed over the different chromosome although it significantly accumulates (p-value < 0.01) in Chr2 (sexual chromosome) and Chr19 (viral resistance chromosome) with p-values 6.68x10^-3^ and 2.06x10^-90^ respectively. 

We determine the gene features occupied by each H3K27me3 mark and the H3K27me3 marked genes. 

```{r k27_peaks_annotation}
library(TxDb.Otauri.JGI)
txdb <- TxDb.Otauri.JGI

ostta.genes <- as.data.frame(genes(txdb))
head(ostta.genes)

target.genes <- vector(mode = "character",length = nrow(ll.20))
peak.annotation <- vector(mode = "character",length = nrow(ll.20))

for(i in 1:nrow(ll.20))
{
 current.chr <- ll.20[i,1]
 current.start <- ll.20[i,2]
 current.end <- ll.20[i,3]

 complete.gene.body <- subset(ostta.genes, 
                              seqnames == current.chr & 
                               start >= current.start & 
                               end <= current.end)
 
 overlap.tss.pos.strand <- subset(ostta.genes, 
                              seqnames == current.chr & 
                              start >= current.start & start <= current.end & 
                              strand == "+")
 
 overlap.tss.neg.strand <- subset(ostta.genes, 
                              seqnames == current.chr & 
                              end >= current.start & end <= current.end & 
                              strand == "-")
 
 overlap.tes.pos.strand <- subset(ostta.genes, 
                              seqnames == current.chr & 
                              end >= current.start & end <= current.end & 
                              strand == "+")
 
 overlap.tes.neg.strand <- subset(ostta.genes, 
                              seqnames == current.chr & 
                              start >= current.start & start <= current.end & 
                              strand == "-")

 inside.gene.body <- subset(ostta.genes, seqnames == current.chr & 
                             start <= current.start & end >= current.end)
 
 
 if(nrow(complete.gene.body) > 0)
 {
  target.genes[i] <- paste(unlist(complete.gene.body$gene_id),collapse = ",")
  peak.annotation[i] <- "gene.body"
 } else if(nrow(overlap.tss.pos.strand) > 0)
 {
  target.genes[i] <- paste(unlist(overlap.tss.pos.strand$gene_id),collapse = ",")
  peak.annotation[i] <- "TSS"
 } else if(nrow(overlap.tss.neg.strand) > 0)
 {
  target.genes[i] <- paste(unlist(overlap.tss.neg.strand$gene_id),collapse = ",")
  peak.annotation[i] <- "TSS"
 } else if(nrow(overlap.tes.pos.strand) > 0)
 {
  target.genes[i] <- paste(unlist(overlap.tes.pos.strand$gene_id),collapse = ",")
  peak.annotation[i] <- "TES"
 } else if(nrow(overlap.tes.neg.strand) > 0)
 {
  target.genes[i] <- paste(unlist(overlap.tes.neg.strand$gene_id),collapse = ",")
  peak.annotation[i] <- "TES"
 } else if(nrow(inside.gene.body) > 0)
 {
  target.genes[i] <- paste(unlist(inside.gene.body$gene_id),collapse = ",")
  peak.annotation[i] <- "internal.gene.body"
 } else
 {
  peak.annotation[i] <- "intergenic"
 }
}

freq.peak.annotation <- table(peak.annotation)
#jpeg(filename = "images/h3k27me3_pie_chart_ll_20.jpg",res = 300,width = 900,height = 900)
#par(mar=c(0,0,0,0))
pie(freq.peak.annotation[c("gene.body",
                           "internal.gene.body",
                           "TSS", 
                           "TES",
                           "intergenic")],
    labels = c("Full Gene Body", 
               "Gene Body", 
               "TSS", 
               "TES", 
               "Intergenic"),
    clockwise = T,border = "black",col=c("blue","cyan","green","red","grey"))
#dev.off()

percentage.freq.peak.annotation <- 100*freq.peak.annotation/sum(freq.peak.annotation)


percentage.freq.peak.annotation["gene.body"] +
percentage.freq.peak.annotation["internal.gene.body"]

percentage.freq.peak.annotation

ll.20.peak.annotation <- cbind(ll.20[,1:3],data.frame(peak.annotation,target.genes))

head(ll.20.peak.annotation)

write.table(x = ll.20.peak.annotation,file = "tables/H3K27me3_LL_20C_peak_annotation.tsv",quote = F,sep = "\t",row.names = F)

all.k27.genes <- (unlist(sapply(X =ll.20.peak.annotation$target.genes,
                                FUN = function(x) {strsplit(x,",")[[1]]})))
names(all.k27.genes) <- NULL
length(all.k27.genes)
length(all.k27.genes)/7668
```
We found 491 genes associated to H3K27me3 marks corresponding to 6.4% of the entire Ostreococcus genome. H3K27me3 is mostly located occupying either the entire gene body (23.75%) or internal regions of gene body (41.86%). Although some regions occupied by H3K27me3 has been found overlapping only with the TSS (11.04%) or the TES (15.42%). H3K27me3 was found in intergenic regions only in 7.92% of the cases. 

Next we use metageneplots to determine the shape of the H3K27me3 over its occupied genes. We consider three different sets of marked genes, genes with their entire gene body marked, genes with H3K27 inside their gene body with no overlap in the TSS or TES and genes with H3K27me3 overlapping their TSS or TES

```{r k27_metageneplots}
## Function to compute ChIP signal over different tiles
ntile.signal <- function(signal.vector,number.tiles)
{
 step.len <- floor(length(signal.vector)/number.tiles)
 number.plus.1 <- length(signal.vector) %% number.tiles
 pos.plus.1 <- sort(sample(1:100,number.plus.1,replace = F))
 
 acum <- c(0)
 for(i in 1:number.tiles)
 {
  if(i %in% pos.plus.1)
  {
   acum <- c(acum,acum[i]+step.len+1)
  } else
  {
   acum <- c(acum,acum[i]+step.len)
  }
 }
 
 acum[1] <- 1
 
 n.tile.signal.vector <- vector(mode = "numeric",length = number.tiles)
 for(i in 2:length(acum))
 {
  n.tile.signal.vector[(i-1)] <- mean(signal.vector[acum[i-1]:acum[i]])
 }
 
 return(n.tile.signal.vector)
}

## Function to generate the data for a metageneplot
metageneplot <- function(gene.names,genes,bw.file.name,ntile.txs,ntile.gene.body)
{
 res <- matrix(nrow=length(gene.names),ncol=2*ntile.txs+ntile.gene.body)
 rownames(res) <- gene.names
 
 bw.file <- import.bw(con = bw.file.name)
 bw.file <- as.data.frame(bw.file)
 
 for(i in 1:length(gene.names))
 {
  #print(i)
  current.gene.chr <- genes[gene.names[i],"seqnames"]
  current.gene.start <- genes[gene.names[i],"start"]
  current.gene.end <- genes[gene.names[i],"end"]
  current.gene.strand <- genes[gene.names[i],"strand"]
  
  gene.body.signal <- subset(bw.file, seqnames == current.gene.chr & 
                              start >= current.gene.start & 
                              end <= current.gene.end)[["score"]]

  upstream.start.signal <- subset(bw.file, seqnames == current.gene.chr & 
                                   start >= (current.gene.start - 2000) & 
                                   end <= current.gene.start)[["score"]]
  
  downstream.end.signal <- subset(bw.file, seqnames == current.gene.chr & 
                                   start >= current.gene.end & 
                                   end <= (current.gene.end + 2000))[["score"]]
  
  ntile.signal.complete <- 
   c(ntile.signal(signal.vector = upstream.start.signal,number.tiles = ntile.txs),
     ntile.signal(signal.vector = gene.body.signal,number.tiles = ntile.gene.body),
     ntile.signal(signal.vector = downstream.end.signal,number.tiles = ntile.txs))
  
  if(current.gene.strand == "+")
  {
   res[i,] <- ntile.signal.complete
  } else
  {
   res[i,] <- rev(ntile.signal.complete)
  }
 }
 
 return(res)
}

## Genes with their entire gene body marked
k27.gene.body <- (unlist(sapply(X = subset(ll.20.peak.annotation, peak.annotation == "gene.body")[["target.genes"]],
                                FUN = function(x) {strsplit(x,",")[[1]]})))
names(k27.gene.body) <- NULL
length(k27.gene.body)

## Genes with H3K27 inside their gene body with no overlap
## in the TSS or TES.
k27.internal.gene.body <- (unlist(sapply(X = subset(ll.20.peak.annotation, peak.annotation == "internal.gene.body")[["target.genes"]],
                                FUN = function(x) {strsplit(x,",")[[1]]})))
names(k27.internal.gene.body) <- NULL
length(k27.internal.gene.body)

## Genes with H3K27me3 overlapping their TSS
k27.tss <- (unlist(sapply(X = subset(ll.20.peak.annotation, peak.annotation == "TSS")[["target.genes"]],
                                         FUN = function(x) {strsplit(x,",")[[1]]})))
names(k27.tss) <- NULL
length(k27.tss)

## Genes with H3K27me3 overlapping their TES
k27.tes <- (unlist(sapply(X = subset(ll.20.peak.annotation, peak.annotation == "TES")[["target.genes"]],
                          FUN = function(x) {strsplit(x,",")[[1]]})))
names(k27.tes) <- NULL
length(k27.tes)
```

Next we generate the metageneplots for genes where the H3K27me3 occupies their entire gene body including TSS and TES. 

```{r k27_metageneplots_gene_body}
library(rtracklayer)
## Genes with their entire gene body marked
k27.gene.body <- (unlist(sapply(X = subset(ll.20.peak.annotation, peak.annotation == "gene.body")[["target.genes"]],
                                FUN = function(x) {strsplit(x,",")[[1]]})))
names(k27.gene.body) <- NULL
length(k27.gene.body)

## Genes with H3K27 inside their gene body with no overlap
## in the TSS or TES.
k27.internal.gene.body <- (unlist(sapply(X = subset(ll.20.peak.annotation, peak.annotation == "internal.gene.body")[["target.genes"]],
                                FUN = function(x) {strsplit(x,",")[[1]]})))
names(k27.internal.gene.body) <- NULL
length(k27.internal.gene.body)

## Genes with H3K27me3 overlapping their TSS
k27.tss <- (unlist(sapply(X = subset(ll.20.peak.annotation, peak.annotation == "TSS")[["target.genes"]],
                                         FUN = function(x) {strsplit(x,",")[[1]]})))
names(k27.tss) <- NULL
length(k27.tss)

## Genes with H3K27me3 overlapping their TES
k27.tes <- (unlist(sapply(X = subset(ll.20.peak.annotation, peak.annotation == "TES")[["target.genes"]],
                          FUN = function(x) {strsplit(x,",")[[1]]})))
names(k27.tes) <- NULL
length(k27.tes)

metageneplot.data.k27.gene.body.1 <- 
 metageneplot(gene.names = k27.gene.body,
              genes = ostta.genes,
              bw.file.name = "chipseq_data/bw_data/h3k27me3/temp20/h3k27me3_20_chip_1.bw",
              ntile.txs = 10,
              ntile.gene.body = 20)
median.metageneplot.data.k27.gene.body.1 <- 
 apply(X = metageneplot.data.k27.gene.body.1,
      MARGIN = 2,
      FUN = median,na.rm = T)

metageneplot.data.k27.gene.body.2 <- 
 metageneplot(gene.names = k27.gene.body,
              genes = ostta.genes,
              bw.file.name = "chipseq_data/bw_data/h3k27me3/temp20/h3k27me3_20_chip_2n.bw",
              ntile.txs = 10,
              ntile.gene.body = 20)
median.metageneplot.data.k27.gene.body.2 <- 
 apply(X = metageneplot.data.k27.gene.body.2,
      MARGIN = 2,
      FUN = median,na.rm = T)

metageneplot.data.k27.gene.body.3 <- 
 metageneplot(gene.names = k27.gene.body,
              genes = ostta.genes,
              bw.file.name = "chipseq_data/bw_data/h3k27me3/temp20/h3k27me3_20_chip_3n.bw",
              ntile.txs = 10,
              ntile.gene.body = 20)
median.metageneplot.data.k27.gene.body.3 <- 
 apply(X = metageneplot.data.k27.gene.body.3,
      MARGIN = 2,
      FUN = median,na.rm = T)

median.metageneplot.data.k27.gene.body <- (median.metageneplot.data.k27.gene.body.1 + median.metageneplot.data.k27.gene.body.2 +                                            median.metageneplot.data.k27.gene.body.3)/3

plot(smooth.spline(x=1:40,y=median.metageneplot.data.k27.gene.body,spar = 0.5),
     type="l",col="blue",lwd=3,ylim=c(0,45),axes=F,xlab="",ylab="")
axis(side = 1, at=c(0,10,30,40),labels = c("-2Kb","TSS","TES","2Kb"),las=2,lwd=2)
axis(side = 2,lwd=2)
```
Next we generate the metageneplots for genes where the H3K27me3 occupies part of the gene body without overlapping with the TSS and TES. 

```{r k27_metageneplots_internal_gene_body}
metageneplot.data.k27.internal.gene.body.1 <- 
 metageneplot(gene.names = k27.internal.gene.body,
              genes = ostta.genes,
              bw.file.name = "chipseq_data/bw_data/h3k27me3/temp20/h3k27me3_20_chip_1.bw",
              ntile.txs = 10,
              ntile.gene.body = 20)
median.metageneplot.data.k27.internal.gene.body.1 <- 
 apply(X = metageneplot.data.k27.internal.gene.body.1,
      MARGIN = 2,
      FUN = median,na.rm = T)

metageneplot.data.k27.internal.gene.body.2 <- 
 metageneplot(gene.names = k27.internal.gene.body,
              genes = ostta.genes,
              bw.file.name = "chipseq_data/bw_data/h3k27me3/temp20/h3k27me3_20_chip_2n.bw",
              ntile.txs = 10,
              ntile.gene.body = 20)
median.metageneplot.data.k27.internal.gene.body.2 <- 
 apply(X = metageneplot.data.k27.internal.gene.body.2,
      MARGIN = 2,
      FUN = median,na.rm = T)

metageneplot.data.k27.internal.gene.body.3 <- 
 metageneplot(gene.names = k27.internal.gene.body,
              genes = ostta.genes,
              bw.file.name = "chipseq_data/bw_data/h3k27me3/temp20/h3k27me3_20_chip_3n.bw",
              ntile.txs = 10,
              ntile.gene.body = 20)
median.metageneplot.data.k27.internal.gene.body.3 <- 
 apply(X = metageneplot.data.k27.internal.gene.body.3,
      MARGIN = 2,
      FUN = median,na.rm = T)

median.metageneplot.data.k27.internal.gene.body <- (median.metageneplot.data.k27.internal.gene.body.1 + 
 median.metageneplot.data.k27.internal.gene.body.2 +                                            median.metageneplot.data.k27.internal.gene.body.3)/3

plot(smooth.spline(x=1:40,y=median.metageneplot.data.k27.internal.gene.body,spar = 0.5),
     type="l",col="cyan",lwd=3,ylim=c(0,10),axes=F,xlab="",ylab="")
axis(side = 1, at=c(0,10,30,40),labels = c("-2Kb","TSS","TES","2Kb"),las=2,lwd=2)
axis(side = 2,lwd=2)
```

Next we generate the metageneplots for genes where the H3K27me3 occupies only the TSS and the beginning of the gene.  

```{r k27_metageneplots_tss}
metageneplot.data.k27.tss.1 <- 
 metageneplot(gene.names = k27.tss,
              genes = ostta.genes,
              bw.file.name = "chipseq_data/bw_data/h3k27me3/temp20/h3k27me3_20_chip_1.bw",
              ntile.txs = 10,
              ntile.gene.body = 20)
median.metageneplot.data.k27.tss.1 <- 
 apply(X = metageneplot.data.k27.tss.1,
      MARGIN = 2,
      FUN = median,na.rm = T)

metageneplot.data.k27.tss.2 <- 
 metageneplot(gene.names = k27.tss,
              genes = ostta.genes,
              bw.file.name = "chipseq_data/bw_data/h3k27me3/temp20/h3k27me3_20_chip_2n.bw",
              ntile.txs = 10,
              ntile.gene.body = 20)
median.metageneplot.data.k27.tss.2 <- 
 apply(X = metageneplot.data.k27.tss.2,
      MARGIN = 2,
      FUN = median,na.rm = T)

metageneplot.data.k27.tss.3 <- 
 metageneplot(gene.names = k27.tss,
              genes = ostta.genes,
              bw.file.name = "chipseq_data/bw_data/h3k27me3/temp20/h3k27me3_20_chip_3n.bw",
              ntile.txs = 10,
              ntile.gene.body = 20)
median.metageneplot.data.k27.tss.3 <- 
 apply(X = metageneplot.data.k27.tss.3,
      MARGIN = 2,
      FUN = median,na.rm = T)



median.metageneplot.data.k27.tss <- 
 (median.metageneplot.data.k27.tss.1 + 
 median.metageneplot.data.k27.tss.2 +                                            median.metageneplot.data.k27.tss.3)/3

plot(smooth.spline(x=1:40,y=median.metageneplot.data.k27.tss,spar = 0.5),
     type="l",col="green",lwd=3,ylim=c(0,20),axes=F,xlab="",ylab="")
axis(side = 1, at=c(0,10,30,40),labels = c("-2Kb","TSS","TES","2Kb"),las=2,lwd=2)
axis(side = 2,lwd=2)
```

Next we generate the metageneplots for genes where the H3K27me3 occupies only the TES and the end of the gene.  

```{r k27_metageneplots_tes}
metageneplot.data.k27.tes.1 <- 
 metageneplot(gene.names = k27.tes,
              genes = ostta.genes,
              bw.file.name = "chipseq_data/bw_data/h3k27me3/temp20/h3k27me3_20_chip_1.bw",
              ntile.txs = 10,
              ntile.gene.body = 20)
median.metageneplot.data.k27.tes.1 <- 
 apply(X = metageneplot.data.k27.tes.1,
      MARGIN = 2,
      FUN = median,na.rm = T)

metageneplot.data.k27.tes.2 <- 
 metageneplot(gene.names = k27.tes,
              genes = ostta.genes,
              bw.file.name = "chipseq_data/bw_data/h3k27me3/temp20/h3k27me3_20_chip_2n.bw",
              ntile.txs = 10,
              ntile.gene.body = 20)
median.metageneplot.data.k27.tes.2 <- 
 apply(X = metageneplot.data.k27.tes.2,
      MARGIN = 2,
      FUN = median,na.rm = T)

metageneplot.data.k27.tes.3 <- 
 metageneplot(gene.names = k27.tes,
              genes = ostta.genes,
              bw.file.name = "chipseq_data/bw_data/h3k27me3/temp20/h3k27me3_20_chip_3n.bw",
              ntile.txs = 10,
              ntile.gene.body = 20)
median.metageneplot.data.k27.tes.3 <- 
 apply(X = metageneplot.data.k27.tes.3,
      MARGIN = 2,
      FUN = median,na.rm = T)



median.metageneplot.data.k27.tes <- 
 (median.metageneplot.data.k27.tes.1 + 
 median.metageneplot.data.k27.tes.2 +                                            median.metageneplot.data.k27.tes.3)/3

plot(smooth.spline(x=1:40,y=median.metageneplot.data.k27.tes,spar = 0.5),
     type="l",col="red",lwd=3,ylim=c(0,20),axes=F,xlab="",ylab="")
axis(side = 1, at=c(0,10,30,40),labels = c("-2Kb","TSS","TES","2Kb"),las=2,lwd=2)
axis(side = 2,lwd=2)
```

In order to compare the different metageneplots are represented in the same plot.

```{r all_k27_metageneplots}
#jpeg(filename = "images/h3k27me3_metageneplots.jpg",res = 250,width = 900,height = 900)
plot(smooth.spline(x=1:40,y=median.metageneplot.data.k27.gene.body,spar = 0.5),
     type="l",col="blue",lwd=3,ylim=c(0,45),axes=F,xlab="",ylab="")
lines(smooth.spline(x=1:40,y=median.metageneplot.data.k27.internal.gene.body,spar = 0.5),
      type="l",col="cyan",lwd=3)
lines(smooth.spline(x=1:40,y=median.metageneplot.data.k27.tss,spar = 0.5),
      type="l",col="green",lwd=3)
lines(smooth.spline(x=1:40,y=median.metageneplot.data.k27.tes,spar = 0.5),
      type="l",col="red",lwd=3)
axis(side = 1, at=c(0,10,30,40),labels = c("-2Kb","TSS","TES","2Kb"),las=2,lwd=2)
axis(side = 2,lwd=2)
#dev.off()
```

It can be observed that the level of H3K27me3 mark varies depending on the part of gene that it occupies. Next we check if these differences are significant.  

```{r metageneplots_significance}
max.metageneplot.data.k27.gene.body.1 <- 
 apply(X = metageneplot.data.k27.gene.body.1,
       MARGIN = 1,
       FUN = max,na.rm = T)
max.metageneplot.data.k27.gene.body.2 <- 
 apply(X = metageneplot.data.k27.gene.body.2,
       MARGIN = 1,
       FUN = max,na.rm = T)
max.metageneplot.data.k27.gene.body.3 <- 
 apply(X = metageneplot.data.k27.gene.body.3,
       MARGIN = 1,
       FUN = max,na.rm = T)

max.metageneplot.data.k27.gene.body <- (max.metageneplot.data.k27.gene.body.1 + 
                                   max.metageneplot.data.k27.gene.body.2 +
                                   max.metageneplot.data.k27.gene.body.3)/3

max.metageneplot.data.k27.internal.gene.body.1 <- 
 apply(X = metageneplot.data.k27.internal.gene.body.1,
       MARGIN = 1,
       FUN = max,na.rm = T)
max.metageneplot.data.k27.internal.gene.body.2 <- 
 apply(X = metageneplot.data.k27.internal.gene.body.2,
       MARGIN = 1,
       FUN = max,na.rm = T)
max.metageneplot.data.k27.internal.gene.body.3 <- 
 apply(X = metageneplot.data.k27.internal.gene.body.3,
       MARGIN = 1,
       FUN = max,na.rm = T)

max.metageneplot.data.k27.internal.gene.body <- (max.metageneplot.data.k27.internal.gene.body.1 + 
                                         max.metageneplot.data.k27.internal.gene.body.2 +
                                         max.metageneplot.data.k27.internal.gene.body.3)/3


max.metageneplot.data.k27.tss.1 <- 
 apply(X = metageneplot.data.k27.tss.1,
       MARGIN = 1,
       FUN = max,na.rm = T)
max.metageneplot.data.k27.tss.2 <- 
 apply(X = metageneplot.data.k27.tss.2,
       MARGIN = 1,
       FUN = max,na.rm = T)
max.metageneplot.data.k27.tss.3 <- 
 apply(X = metageneplot.data.k27.tss.3,
       MARGIN = 1,
       FUN = max,na.rm = T)
max.metageneplot.data.k27.tss <- (max.metageneplot.data.k27.tss.1 + 
                                   max.metageneplot.data.k27.tss.2 +
                                   max.metageneplot.data.k27.tss.3)/3

max.metageneplot.data.k27.tes.1 <- 
 apply(X = metageneplot.data.k27.tes.1,
       MARGIN = 1,
       FUN = max,na.rm = T)
max.metageneplot.data.k27.tes.2 <- 
 apply(X = metageneplot.data.k27.tes.2,
       MARGIN = 1,
       FUN = max,na.rm = T)
max.metageneplot.data.k27.tes.3 <- 
 apply(X = metageneplot.data.k27.tes.3,
       MARGIN = 1,
       FUN = max,na.rm = T)
max.metageneplot.data.k27.tes <- (max.metageneplot.data.k27.tes.1 + 
                                  max.metageneplot.data.k27.tes.2 +
                                  max.metageneplot.data.k27.tes.3)/3

shapiro.test(max.metageneplot.data.k27.gene.body)
shapiro.test(max.metageneplot.data.k27.internal.gene.body)
shapiro.test(max.metageneplot.data.k27.tss)
shapiro.test(max.metageneplot.data.k27.tes)

wilcox.test(x = max.metageneplot.data.k27.gene.body,
            y = max.metageneplot.data.k27.internal.gene.body,
            alternative = "greater")$p.value
wilcox.test(x = max.metageneplot.data.k27.gene.body,
            y = max.metageneplot.data.k27.tss,
            alternative = "greater")
wilcox.test(x = max.metageneplot.data.k27.gene.body,
            y = max.metageneplot.data.k27.tes,
            alternative = "greater")
wilcox.test(x = max.metageneplot.data.k27.tss,
            y = max.metageneplot.data.k27.tes)
wilcox.test(x = max.metageneplot.data.k27.tss,
            y = max.metageneplot.data.k27.tes)
```

Indeed the differences between the metageneplots are highly significant except the ones corresponding to the marks overlapping with the TSS or TES. 

Now we analyse the gene expression level of marked genes.

```{r h3k27me3_gene_expression}
temp.gene.expression <- read.table(file="rnaseq_data/ostta_temperature_gene_expression.tsv",header = T,sep = "\t")

gene.expression.20 <- 
 (temp.gene.expression$T20_1 + 
  temp.gene.expression$T20_2 + 
  temp.gene.expression$T20_3)/3

names(gene.expression.20) <- rownames(temp.gene.expression)

non.k27.genes <- setdiff(names(gene.expression.20),c(k27.gene.body,k27.tss,k27.tes,k27.internal.gene.body))

summary(gene.expression.20[k27.gene.body])
summary(gene.expression.20[k27.internal.gene.body])
summary(gene.expression.20[k27.tss])
summary(gene.expression.20[k27.tes])
summary(gene.expression.20[non.k27.genes])

#jpeg(filename = "images/h3k27me3_expression_barplot_different_genes.jpg",res = 250,width = 600,height = 1000)
boxplot(gene.expression.20[k27.gene.body],
        gene.expression.20[k27.tss],
        gene.expression.20[k27.internal.gene.body],
        gene.expression.20[k27.tes],
        gene.expression.20[non.k27.genes],
        col=c("blue","green","cyan","red","grey"),
        outline=F)
#dev.off()

wilcox.test(x = gene.expression.20[k27.gene.body], 
            y = gene.expression.20[k27.tss],
            alternative = "less")
wilcox.test(x = gene.expression.20[k27.tss], 
            y = gene.expression.20[k27.internal.gene.body],
            alternative = "less")

wilcox.test(x = gene.expression.20[k27.internal.gene.body],
            y = gene.expression.20[k27.tes],
            alternative = "less")

wilcox.test(x = gene.expression.20[k27.tes],
            y = gene.expression.20[non.k27.genes],
            alternative = "less")
```

We can see all H3K27me3 genes present levels of expression significantly lower than non-marked genes. Specifically, genes whose entire gene body or TSS is occupied by H3K27me3 are completely repressed. Whereas, genes where only the TES or gene body is occupied by H3K27me3 are expressed although at a significantly lower level than non-marked genes.  

In what follows we perform a functional analysis over the H3K27me3 marked genes. First, we analyze the relationship between H3K27me3 peaks and transposable elements. 

```{r transposable_elements}
dna.transposons <- read.table(file="te_analysis/dna_transposons_unsorted.bed",sep="\t",header=F)
nrow(dna.transposons)

ltr.transposons <- read.table(file="te_analysis/ltr_transposons_unsorted.bed",sep="\t", header=F)
nrow(ltr.transposons)

unclassified.transposons <- read.table(file="te_analysis/unknown_transposons_unsorted.bed", sep="\t",header = F)
nrow(unclassified.transposons)

overlapping.peaks <- function(target.peak, peaks)
{
 subset(peaks, V1 == target.peak[1] & (
             (V2 >= target.peak[2] & V2 <= target.peak[3]) |
             (V3 >= target.peak[2] & V3 <= target.peak[3]) |
             (V2 <= target.peak[2] & V3 >= target.peak[3])))
}


peak.te <- vector(mode = "character",length = nrow(ll.20.peak.annotation))

for(i in 1:nrow(ll.20.peak.annotation))
{
 current.peak <- unlist(ll.20.peak.annotation[i,1:3])
 
 if(nrow(overlapping.peaks(current.peak,dna.transposons)) > 0)
 {
  peak.te[i] <- "DNA_transposon"
 } else if (nrow(overlapping.peaks(current.peak,ltr.transposons)) > 0)
 {
  peak.te[i] <- "LTR_transposon"
 } else if (nrow(overlapping.peaks(current.peak,unclassified.transposons)) > 0)
 {
  peak.te[i] <- "Unclassified_transposon"
 } else
 {
  peak.te[i] <- "No_transposon"
 }
}

100*table(peak.te)/sum(table(peak.te))

pie(table(peak.te), col=c("black","grey","white","lightblue"))

nrow(dna.transposons)
100*table(peak.te)["DNA_transposon"]/nrow(dna.transposons)

nrow(ltr.transposons)
100*table(peak.te)["LTR_transposon"]/nrow(ltr.transposons)

nrow(unclassified.transposons)
100*table(peak.te)["Unclassified_transposon"]/nrow(unclassified.transposons)


table(ll.20.peak.annotation$peak.annotation[peak.te == "DNA_transposon"])
table(ll.20.peak.annotation$peak.annotation[peak.te == "LTR_transposon"])
table(ll.20.peak.annotation$peak.annotation[peak.te == "Unclassified_transposon"])

table(ll.20.peak.annotation$peak.annotation)

(table(ll.20.peak.annotation$peak.annotation[peak.te == "DNA_transposon"])["intergenic"] +
table(ll.20.peak.annotation$peak.annotation[peak.te == "LTR_transposon"])["intergenic"] +
table(ll.20.peak.annotation$peak.annotation[peak.te == "Unclassified_transposon"])["intergenic"]) / table(ll.20.peak.annotation$peak.annotation)["intergenic"]

ll.20.peak.annotation <- cbind(ll.20.peak.annotation,peak.te)
write.table(x = ll.20.peak.annotation,file = "tables/H3K27me3_LL_20C_peak_annotation.tsv",quote = F,sep = "\t",row.names = F)
```

H3K27me3 is only partially associated with Transposable Elements. More than 81% of the peaks did not overlap with any TE. Only 1.5%, 0.6% and 17.3% of the H3K27me3 peaks were found overlapping with DNA, LTR and unclassified transposons respectively. Most TE were neither associated with H3K27me3 peaks, specifically 7 out of 35 (20%) DNA transposons, 3 out of 69 (4.3%) LTR transposons and 83 out of 988 unclassified transposons were H3K27me3 marked. Most H3K27me3 marks located at intergenic regions (81.6%) were associated with transposons    


```{r k27_go_terms}
library(clusterProfiler)
library(enrichplot)
library(org.Otauriv5.eg.db)
k27.enrichment <- enrichGO(gene = all.k27.genes, pvalueCutoff = 0.05,qvalueCutoff = 0.05,
                       OrgDb = org.Otauriv5.eg.db,
                       ont = "BP", 
                       keyType = "GID")
k27.enrichment.df <- as.data.frame(k27.enrichment)

k27.go.terms <- enrichGO(gene = all.k27.genes, pvalueCutoff = 1,qvalueCutoff = 1,
                       OrgDb = org.Otauriv5.eg.db,
                       ont = "BP", 
                       keyType = "GID")
k27.go.terms.df <- as.data.frame(k27.go.terms)


write.table(x = k27.go.terms.df, file = "go_enrichment_k27.tsv",quote = F,sep = "\t",row.names = F)

treeplot(x = pairwise_termsim(k27.go.terms))

#jpeg(filename = "images/h3k27me3_revigo_treemap.jpg",res = 300,width = 1400,height = 900)

library(treemap)

revigo.names <- c("term_ID","description","frequency","value","uniqueness","dispensability","representative");
revigo.data <- rbind(c("GO:0000003","reproduction",1.1154428097959028,0.35455656429893245,1,-0,"reproduction"),
                     c("GO:0001887","selenium compound metabolic process",0.0031524026730797234,0.33700103077893184,0.9533150770650385,0.05458229,"selenium compound metabolic process"),
                     c("GO:0002376","immune system process",0.9427113541805001,0.3647586988245658,1,-0,"immune system process"),
                     c("GO:0006099","tricarboxylic acid cycle",0.5040640114710979,0.40392617241762396,0.926776665398285,0.08410724,"tricarboxylic acid cycle"),
                     c("GO:0006103","2-oxoglutarate metabolic process",0.017201421622692253,1.0244715302516616,0.900036020673125,0.06139517,"2-oxoglutarate metabolic process"),
                     c("GO:0006525","arginine metabolic process",0.425056765430777,0.6822743332528265,0.8414193966872098,0.31651609,"2-oxoglutarate metabolic process"),
                     c("GO:0006526","arginine biosynthetic process",0.2840884163423315,0.6822743332528265,0.8197776139390401,0.55006631,"2-oxoglutarate metabolic process"),
                     c("GO:0006631","fatty acid metabolic process",1.936945636803579,0.661709121736741,0.8363267566648038,0.47549756,"2-oxoglutarate metabolic process"),
                     c("GO:0019344","cysteine biosynthetic process",0.13937119746048948,0.3647586988245658,0.8182547791420745,0.64560862,"2-oxoglutarate metabolic process"),
                     c("GO:0019541","propionate metabolic process",0.06516280052453614,0.3121779710446768,0.8755895054022355,0.56447247,"2-oxoglutarate metabolic process"),
                     c("GO:0030258","lipid modification",0.7621839254181456,0.3121779710446768,0.9072895155224819,0.63919224,"2-oxoglutarate metabolic process"),
                     c("GO:0042128","nitrate assimilation",0.06915814433459262,0.39602842215015766,0.872968299057155,0.3471953,"2-oxoglutarate metabolic process"),
                     c("GO:0043648","dicarboxylic acid metabolic process",1.2968131796952338,0.34085777107681914,0.8591201035057788,0.53948233,"2-oxoglutarate metabolic process"),
                     c("GO:0046487","glyoxylate metabolic process",0.1082932886998466,0.42431833266535085,0.8788025002701066,0.59108986,"2-oxoglutarate metabolic process"),
                     c("GO:0007017","microtubule-based process",1.4522840599239462,2.9055417594493695,0.9846414615040694,0.01131391,"microtubule-based process"),
                     c("GO:0007018","microtubule-based movement",0.6202173565622278,3.3618830741865118,0.9760729714791799,0,"microtubule-based movement"),
                     c("GO:0007049","cell cycle",2.8073119376140743,1.055588520937294,0.9834821578378161,0.01350265,"cell cycle"),
                     c("GO:0007154","cell communication",9.671968224175414,0.47456381848456913,0.9807530444684127,0.0178355,"cell communication"),
                     c("GO:0009267","cellular response to starvation",0.27913921808780134,0.9270279660356058,0.8817472625874171,0.00961211,"cellular response to starvation"),
                     c("GO:0000160","phosphorelay signal transduction system",2.0739975137672992,0.6100474021856732,0.8687040167686203,0.43251558,"cellular response to starvation"),
                     c("GO:0007166","cell surface receptor signaling pathway",1.7699816731780171,0.5778037246528189,0.8709358661055234,0.56185435,"cellular response to starvation"),
                     c("GO:0071496","cellular response to external stimulus",0.6305470826004429,0.7229162230356428,0.9204539188066697,0.65860564,"cellular response to starvation"),
                     c("GO:0015074","DNA integration",0.7252350055806647,2.657762206770456,0.8035461550887487,0.01150715,"DNA integration"),
                     c("GO:0000271","polysaccharide biosynthetic process",0.8445801891875397,0.42431833266535085,0.8524365468053611,0.2912699,"DNA integration"),
                     c("GO:0005975","carbohydrate metabolic process",5.3400641443698875,0.6489525204320686,0.9194445968125812,0.10937561,"DNA integration"),
                     c("GO:0006259","DNA metabolic process",5.6268046211314795,2.0741453286090206,0.7883371559851247,0.44302427,"DNA integration"),
                     c("GO:0006270","DNA replication initiation",0.18677061560434888,1.524614033979224,0.8223941766176285,0.48453834,"DNA integration"),
                     c("GO:0006281","DNA repair",2.895534847137308,0.7716708877524712,0.7159639044019359,0.64562006,"DNA integration"),
                     c("GO:0006298","mismatch repair",0.21782141222315915,0.6187036479778791,0.7758261295536709,0.68512103,"DNA integration"),
                     c("GO:0006468","protein phosphorylation",1.0150465485898383,2.0478177071574097,0.8059994933400036,0.18117637,"DNA integration"),
                     c("GO:0006486","protein glycosylation",0.7526798873357411,0.6623793059324244,0.7700801528471054,0.66477721,"DNA integration"),
                     c("GO:0006508","proteolysis",5.2622572267907,2.215841795989129,0.8331839915576025,0.43614593,"DNA integration"),
                     c("GO:0006637","acyl-CoA metabolic process",0.5539971825062208,0.39602842215015766,0.766633541793333,0.51243251,"DNA integration"),
                     c("GO:0008213","protein alkylation",0.000579213939150692,0.33320452706495884,0.8948472249778151,0.37850183,"DNA integration"),
                     c("GO:0009100","glycoprotein metabolic process",0.9570906483124367,0.6014649273302938,0.8398249210805684,0.35568771,"DNA integration"),
                     c("GO:0009119","ribonucleoside metabolic process",0.3506733719429951,0.39602842215015766,0.7958005207549018,0.45483879,"DNA integration"),
                     c("GO:0009226","nucleotide-sugar biosynthetic process",0.35550672740969513,0.3033112734255381,0.7919369046997367,0.48997503,"DNA integration"),
                     c("GO:0010501","RNA secondary structure unwinding",0.0022996025754365773,0.33700103077893184,0.8841308449932884,0.22153339,"DNA integration"),
                     c("GO:0016266","O-glycan processing",0.009247705105078281,0.6822743332528265,0.832553230099121,0.68426048,"DNA integration"),
                     c("GO:0016310","phosphorylation",5.235381700014107,0.9200287559031499,0.8995173712742447,0.51243059,"DNA integration"),
                     c("GO:0016579","protein deubiquitination",0.0016070105886223452,1.363169883748535,0.8876399511853289,0.40321557,"DNA integration"),
                     c("GO:0018202","peptidyl-histidine modification",1.0281836136778093,0.9270279660356058,0.8252111825191477,0.68739488,"DNA integration"),
                     c("GO:0018342","protein prenylation",0.00018239077232830298,0.3647586988245658,0.8956376571084298,0.35393725,"DNA integration"),
                     c("GO:0019439","aromatic compound catabolic process",2.2395984760807317,0.488426217953643,0.8298501364385389,0.67803842,"DNA integration"),
                     c("GO:0032543","mitochondrial translation",0.2221051306327077,0.7671257790005732,0.8301308068014243,0.30712983,"DNA integration"),
                     c("GO:0034404","nucleobase-containing small molecule biosynthetic process",0.2580336480412168,0.3647586988245658,0.8075679421637405,0.47645433,"DNA integration"),
                     c("GO:0034655","nucleobase-containing compound catabolic process",1.6564680648053147,0.5761569684631386,0.8082436737396426,0.50011278,"DNA integration"),
                     c("GO:0035383","thioester metabolic process",0.5539971825062208,0.39602842215015766,0.9156152717960674,0.69278956,"DNA integration"),
                     c("GO:0042398","cellular modified amino acid biosynthetic process",0.5343877097407617,0.40392617241762396,0.8708997557571673,0.19203267,"DNA integration"),
                     c("GO:0044270","cellular nitrogen compound catabolic process",2.004743244566209,0.4588180186234574,0.8286580702160516,0.69482359,"DNA integration"),
                     c("GO:0046434","organophosphate catabolic process",0.9710632603168847,0.868457312023167,0.8769887779688311,0.41091092,"DNA integration"),
                     c("GO:0046700","heterocycle catabolic process",2.042108705491845,0.4685002750565224,0.8310802481954842,0.69649215,"DNA integration"),
                     c("GO:0070646","protein modification by small protein removal",0.0641152859537317,1.1763933331577958,0.8585287404338742,0.52767687,"DNA integration"),
                     c("GO:0071897","DNA biosynthetic process",0.7667855953091854,2.0080982132635268,0.7764795050343053,0.55438786,"DNA integration"),
                     c("GO:0090304","nucleic acid metabolic process",13.541109943481535,0.5761569684631386,0.7844903754380635,0.55585571,"DNA integration"),
                     c("GO:0140053","mitochondrial gene expression",0.3272164397774754,0.7229162230356428,0.8689292553283973,0.16191549,"DNA integration"),
                     c("GO:1901361","organic cyclic compound catabolic process",2.407836710372425,0.4588180186234574,0.8509596622336734,0.55813667,"DNA integration"),
                     c("GO:0016441","regulation of gene expression",0.049918382594123255,0.7229162230356428,0.9152721684520655,0.09459438,"regulation of gene expression"),
                     c("GO:0006450","regulation of translational fidelity",0.29561600610151356,0.3647586988245658,0.9565840743545636,0.13549893,"regulation of gene expression"),
                     c("GO:0010608","post-transcriptional regulation of gene expression",1.6811968028967992,0.5129770046876073,0.9187967357951126,0.29038639,"regulation of gene expression"),
                     c("GO:0010629","negative regulation of gene expression",1.1219250764340574,0.531092579249663,0.891938194299043,0.58730682,"regulation of gene expression"),
                     c("GO:0017148","negative regulation of translation",0.21779183534115995,0.33700103077893184,0.9025830172260265,0.66877869,"regulation of gene expression"),
                     c("GO:0031326","regulation of cellular biosynthetic process",14.068660464000152,0.3894214410778525,0.8933198148321295,0.55551169,"regulation of gene expression"),
                     c("GO:0033044","regulation of chromosome organization",0.18193972487781543,0.3121779710446768,0.9545386992795665,0.15097414,"regulation of gene expression"),
                     c("GO:0034249","negative regulation of amide metabolic process",0.23996217313971518,0.33700103077893184,0.9072896668585831,0.68743154,"regulation of gene expression"),
                     c("GO:1903311","regulation of mRNA metabolic process",0.4779772015478174,0.3647586988245658,0.9285627652482316,0.34495101,"regulation of gene expression"),
                     c("GO:2000241","regulation of reproductive process",0.1639520511419782,0.3121779710446768,0.9569481717332748,0.14866641,"regulation of gene expression"),
                     c("GO:0030029","actin filament-based process",0.8246601591610891,0.6822743332528265,0.9855133147715953,0.01167087,"actin filament-based process"),
                     c("GO:0030154","cell differentiation",2.279586420543629,0.4717135495803034,0.9548569791025598,0.01433045,"cell differentiation"),
                     c("GO:0009888","tissue development",0.6842315881691289,0.3494179598129715,0.9723962948998165,0.69552749,"cell differentiation"),
                     c("GO:0048229","gametophyte development",0.017253181166190824,0.3647586988245658,0.9784815653260818,0.48986076,"cell differentiation"),
                     c("GO:0048507","meristem development",0.03934957675974807,0.33700103077893184,0.9773200335076074,0.52461364,"cell differentiation"),
                     c("GO:0040011","locomotion",0.5148398554794674,2.9956162519741056,1,-0,"locomotion"),
                     c("GO:0043547","positive regulation of GTPase activity",0.0008897712001421268,0.9270279660356058,0.945105760674195,-0,"positive regulation of GTPase activity"),
                     c("GO:0051168","nuclear export",0.36225518598584233,0.8822380235820269,0.9193145237527721,0.00984617,"nuclear export"),
                     c("GO:0006403","RNA localization",0.38503431460554743,0.5719097997394266,0.9492853307589769,0.48954515,"nuclear export"),
                     c("GO:0006614","SRP-dependent cotranslational protein targeting to membrane",0.1731775735855571,0.3647586988245658,0.9239488841068346,0.45707974,"nuclear export"),
                     c("GO:0006890","retrograde vesicle-mediated transport, Golgi to endoplasmic reticulum",0.1396694310206479,0.3121779710446768,0.9533956800814197,0.22525423,"nuclear export"),
                     c("GO:0015931","nucleobase-containing compound transport",0.5729682875685309,0.3245363289714383,0.953096561853417,0.48659654,"nuclear export"),
                     c("GO:0031503","protein-containing complex localization",0.2928998624379218,0.3647586988245658,0.9641396486534709,0.22721235,"nuclear export"),
                     c("GO:0033750","ribosome localization",0.08453565823400183,0.39602842215015766,0.9615600081516522,0.20497386,"nuclear export"),
                     c("GO:0050658","RNA transport",0.3380539022900098,0.6187036479778791,0.9329722181898726,0.24285839,"nuclear export"),
                     c("GO:0051169","nuclear transport",0.5733749696960196,0.5066832555406448,0.9263142426302791,0.63305059,"nuclear export"),
                     c("GO:0051301","cell division",1.5693197819947182,0.5659276631444706,0.9845137269828846,0.01363807,"cell division"),
                     c("GO:0061982","meiosis I cell cycle process",0.22371953544182982,1.4436467858208988,0.8829935432772962,0.00942192,"meiosis I cell cycle process"),
                     c("GO:0098813","nuclear chromosome segregation",0.33031461816688995,1.0576395600847068,0.9027750717677788,0.64747101,"meiosis I cell cycle process"),
                     c("GO:0070085","glycosylation",0.8372204750500761,0.5642517085182833,0.9603127258883271,0.04818597,"glycosylation"),
                     c("GO:0071103","plasma membrane bounded cell projection organization",0.8233982121957907,2.48892688143064,0.8072006030847753,0.01166889,"plasma membrane bounded cell projection organization"),
                     c("GO:0000027","ribosomal large subunit assembly",0.012821578346646382,0.8154250425427645,0.8458879193561408,0.48240062,"plasma membrane bounded cell projection organization"),
                     c("GO:0000245","spliceosomal complex assembly",0.1148766096848317,0.33700103077893184,0.7176266579933399,0.64218601,"plasma membrane bounded cell projection organization"),
                     c("GO:0000280","nuclear division",0.45164391760787703,0.4717135495803034,0.8236095327076562,0.491872,"plasma membrane bounded cell projection organization"),
                     c("GO:0006415","translational termination",0.15480047090339727,0.3647586988245658,0.7363259627235162,0.48918433,"plasma membrane bounded cell projection organization"),
                     c("GO:0006479","protein methylation",0.000579213939150692,0.33320452706495884,0.8923844376711241,0.51387372,"plasma membrane bounded cell projection organization"),
                     c("GO:0006996","organelle organization",6.781836087483291,0.901501456180904,0.8241728840064373,0.52336479,"plasma membrane bounded cell projection organization"),
                     c("GO:0007010","cytoskeleton organization",1.98151060375585,1.5080424402092374,0.8150504345607171,0.61927091,"plasma membrane bounded cell projection organization"),
                     c("GO:0030030","cell projection organization",1.3282977705833654,1.6133625971416043,0.8512620269157171,0.41569968,"plasma membrane bounded cell projection organization"),
                     c("GO:0042255","ribosome assembly",0.2624923630025938,0.42431833266535085,0.8171364501582283,0.65665084,"plasma membrane bounded cell projection organization"),
                     c("GO:0043248","proteasome assembly",0.04973352708162835,0.39602842215015766,0.8611074492352145,0.52931974,"plasma membrane bounded cell projection organization"),
                     c("GO:0048285","organelle fission",0.5487596096521986,0.30148135432836,0.8356875754692098,0.50144746,"plasma membrane bounded cell projection organization"),
                     c("GO:0051258","protein polymerization",0.1780454354145895,1.0725168728382306,0.8473253852004743,0.32972263,"plasma membrane bounded cell projection organization"),
                     c("GO:0051276","chromosome organization",1.5299726699751368,1.9364166059501378,0.8195931631949569,0.55870698,"plasma membrane bounded cell projection organization"),
                     c("GO:0070475","rRNA base methylation",0.22436283262531206,0.33700103077893184,0.6953574452494768,0.4998146,"plasma membrane bounded cell projection organization"),
                     c("GO:0070897","transcription preinitiation complex assembly",0.11026754557329214,0.39602842215015766,0.7195583694392909,0.56139216,"plasma membrane bounded cell projection organization"),
                     c("GO:0070925","organelle assembly",1.3561320812847646,0.6413546437511954,0.8010719571331075,0.5915788,"plasma membrane bounded cell projection organization"),
                     c("GO:0097435","supramolecular fiber organization",0.801568008540226,0.5477838551470221,0.8581207014692784,0.38125869,"plasma membrane bounded cell projection organization"),
                     c("GO:0120036","plasma membrane bounded cell projection organization",1.1030821378604103,1.7112755303804414,0.8324651987784939,0.3943391,"plasma membrane bounded cell projection organization"),
                     c("GO:0097354","prenylation",0.009942761832059112,0.3647586988245658,0.9497005745047677,0.05901557,"prenylation"),
                     c("GO:2001057","reactive nitrogen species metabolic process",0.11012952012396263,0.33700103077893184,0.9373241093491478,0.09288199,"reactive nitrogen species metabolic process"));

stuff <- data.frame(revigo.data);
names(stuff) <- revigo.names;

stuff$value <- as.numeric( as.character(stuff$value) );
stuff$frequency <- as.numeric( as.character(stuff$frequency) );
stuff$uniqueness <- as.numeric( as.character(stuff$uniqueness) );
stuff$dispensability <- as.numeric( as.character(stuff$dispensability) );


treemap(
 stuff,
 index = c("representative","description"),
 vSize = "value",
 type = "categorical",
 vColor = "representative",
 title = "",
 inflate.labels = TRUE,
 lowerbound.cex.labels = 0,
 position.legend = "none"
)

#dev.off()
```


We next explore the association between ChIP signal and gene expression.

```{r chip_signal_vs_expression}
count.1 <- read.table(file = "chipseq_data/bed_data/h3k27me3/temp20c/k27_20_ll_chip_1_counts.bed",header = F,sep = "\t")
count.2 <- read.table(file="chipseq_data/bed_data/h3k27me3/temp20c/k27_20_ll_chip_2n_counts.bed",header = F,sep = "\t")
count.3 <- read.table(file="chipseq_data/bed_data/h3k27me3/temp20c/k27_20_ll_chip_3n_counts.bed",header = F,sep = "\t")
count.1 <- count.1[,c(1:3,10)]
count.2 <- count.2[,c(1:3,10)]
count.3 <- count.3[,c(1:3,10)]

mapped.reads.k27.20c.1 <- 3146662/1E6 
mapped.reads.k27.20c.2 <- 5697849/1E6
mapped.reads.k27.20c.3 <- 5157582/1E6

rpm.1 <- count.1
rpm.2 <- count.2
rpm.3 <- count.3

rpm.1[,4] <- rpm.1[,4]/mapped.reads.k27.20c.1
rpm.2[,4] <- rpm.2[,4]/mapped.reads.k27.20c.2
rpm.3[,4] <- rpm.3[,4]/mapped.reads.k27.20c.3


plot(log2(rpm.1[,4]),log2(rpm.2[,4]))
plot(log2(rpm.1[,4]),log2(rpm.3[,4]))
plot(log2(rpm.2[,4]),log2(rpm.3[,4]))

rpm <- rpm.1
rpm[,4] <- (rpm.1[,4] + rpm.2[,4] + rpm.3[,4])/3
```

## H3K4me3 genome wide distribution analysis

We load the H3K4me3 peaks identified in cultures grown under constant light at 20ºC. We represent each chromosome using a rectangle. Each H3k4me3 peak is represented in its precise genome location using a dark orange rectangle. 


```{r visualization of H3K4me3 genome wide distribution}
k4.ll.20 <- read.table(file="chipseq_data/bed_data/h3k4me3/h3k4me3_ll_20.bed",
                       header = F)

k4.ll.20 <- k4.ll.20[,1:3]

nrow(k4.ll.20)

#jpeg(filename = "images/h3k27me3_genome_wide_distribution_ll_20.jpg",res = 300,width = 900,height = 900)
#par(mar=c(0,0,0,0))
plot(x=c(0,max.chr.len), y=c(0,4*number.chrs), col = "white", xlab = "", ylab = "", axes=F)  

for(i in 1:number.chrs)
{
 polygon(x = c(0, chr.lens[i], chr.lens[i], 0),   
         y = c(4*i+1, 4*i+1, 4*i-1, 4*i-1),    
         col = "white") 
}

for(i in 1:nrow(k4.ll.20))
{
 current.chr <- k4.ll.20[i,1]
 current.start <- k4.ll.20[i,2]
 current.end <- k4.ll.20[i,3]

 current.line <- (20:1)[current.chr]
 
 polygon(x = c(current.start, current.end, current.end, current.start),   
         y = c(4*current.line+1, 4*current.line+1, 4*current.line-1, 4*current.line-1),    
         col = "orange",border="orange")
}
#dev.off()
```
We identified 773 genomic regions significantly occupied by H3K4me3 in the three different independent biological replicates. We compute the percentage of the whole genome occupied by H3K4me3 and the percentage in each chromosome. A barplot is used to represent the percentage of H3K4me3 in each chromosome. Z-scores and p-values are computed for each chromosome. A gradient from white to dark orange is used to represent significance levels.  

```{r k4_in_chrs}
h3k4me3.peaks.len <- k4.ll.20$V3 - k4.ll.20$V2
sum(h3k4me3.peaks.len)/sum(chr.lens)
k4.ll.20 <- cbind(k4.ll.20,h3k4me3.peaks.len)
colnames(k4.ll.20) <- c("chr","start","end","length")
head(k4.ll.20)
number.of.h3k4.peaks.chr <- vector(mode = "numeric",length = 20)
k4.peaks.len.chr <- vector(mode = "numeric",length = 20)

for(i in 1:20)
{
 number.of.h3k4.peaks.chr[i] <- nrow(subset(k4.ll.20, chr == i))
 k4.peaks.len.chr[i] <- sum(subset(k4.ll.20, chr == i)[,4]) 
}

percentage.chr.k4 <- k4.peaks.len.chr/rev(chr.lens)

k4.z.scores <- (percentage.chr.k4 - mean(percentage.chr.k4))/sd(percentage.chr.k4)

which(k4.z.scores > 2.5)
which(k4.z.scores < -2.5)

p.vals <- pnorm((percentage.chr.k4 - mean(percentage.chr.k4[-20]))/sd(percentage.chr.k4[-20]),lower.tail = F)

p.vals

values.for.colors <- round(-log10(p.vals)*40+1)
values.for.colors[20] <- 100

white.darkorange.palette <- colorRampPalette(c('white','darkorange'))

#jpeg(filename = "images/h3k27me3_chr_barplot_ll_20.jpg",res = 300,width = 1800,height = 900)
barplot(100*percentage.chr.k4,border="black",names.arg = paste0("Chr",1:20),las=2,col=white.darkorange.palette(100)[values.for.colors]
)
abline(h = 100*mean(percentage.chr.k4[-20]),lty=5)
#dev.off()
```
Approximately 3.7% of the Ostreococcus genome is occupied by H3K4me3. 
H3K4me3 is mostly uniformly distributed over the different chromosome although it significantly accumulates (p-value < 0.01) in Chr20 with p-value 5.32x10^-4^.

We determine the gene features occupied by each H3K4me3 mark and the H3K4me3 marked genes. 

```{r k4_peaks_annotation}
library(TxDb.Otauri.JGI)
txdb <- TxDb.Otauri.JGI

ostta.genes <- as.data.frame(genes(txdb))
head(ostta.genes)

target.genes <- vector(mode = "character",length = nrow(k4.ll.20))
peak.annotation <- vector(mode = "character",length = nrow(k4.ll.20))

for(i in 1:nrow(k4.ll.20))
{
 current.chr <- k4.ll.20[i,1]
 current.start <- k4.ll.20[i,2]
 current.end <- k4.ll.20[i,3]
 
 complete.gene.body <- subset(ostta.genes, 
                              seqnames == current.chr & 
                               start >= current.start & 
                               end <= current.end)
 
 overlap.tss.pos.strand <- subset(ostta.genes, 
                                  seqnames == current.chr & 
                                   start >= current.start & start <= current.end & 
                                   strand == "+")
 
 overlap.tss.neg.strand <- subset(ostta.genes, 
                                  seqnames == current.chr & 
                                   end >= current.start & end <= current.end & 
                                   strand == "-")
 
 overlap.tes.pos.strand <- subset(ostta.genes, 
                                  seqnames == current.chr & 
                                   end >= current.start & end <= current.end & 
                                   strand == "+")
 
 overlap.tes.neg.strand <- subset(ostta.genes, 
                                  seqnames == current.chr & 
                                   start >= current.start & start <= current.end & 
                                   strand == "-")
 
 inside.gene.body <- subset(ostta.genes, seqnames == current.chr & 
                             start <= current.start & end >= current.end)
 
 
 if(nrow(complete.gene.body) > 0)
 {
  target.genes[i] <- paste(unlist(complete.gene.body$gene_id),collapse = ",")
  peak.annotation[i] <- "gene.body"
 } else if((nrow(overlap.tss.pos.strand) > 0) | (nrow(overlap.tss.neg.strand) > 0))
 {
  target.genes[i] <- paste(c(unlist(overlap.tss.pos.strand$gene_id),
                             unlist(overlap.tss.neg.strand$gene_id)),collapse = ",")
  peak.annotation[i] <- "TSS"
 } else if(nrow(overlap.tes.pos.strand) > 0)
 {
  target.genes[i] <- paste(unlist(overlap.tes.pos.strand$gene_id),collapse = ",")
  peak.annotation[i] <- "TES"
 } else if(nrow(overlap.tes.neg.strand) > 0)
 {
  target.genes[i] <- paste(unlist(overlap.tes.neg.strand$gene_id),collapse = ",")
  peak.annotation[i] <- "TES"
 } else if(nrow(inside.gene.body) > 0)
 {
  target.genes[i] <- paste(unlist(inside.gene.body$gene_id),collapse = ",")
  peak.annotation[i] <- "internal.gene.body"
 } else
 {
  peak.annotation[i] <- "intergenic"
 }
}

freq.peak.annotation <- table(peak.annotation)
#jpeg(filename = "images/h3k4me3_pie_chart_ll_20.jpg",res = 300,width = 900,height = 900)
#par(mar=c(0,0,0,0))
pie(freq.peak.annotation[c("gene.body",
                           "internal.gene.body",
                           "TSS", 
                           "TES",
                           "intergenic")],
    labels = c("Full Gene Body", 
               "Gene Body", 
               "TSS", 
               "TES", 
               "Intergenic"),
    clockwise = T,border = "black",col=c("blue","cyan","green","red","grey"))
#dev.off()

percentage.freq.peak.annotation <- 100*freq.peak.annotation/sum(freq.peak.annotation)

percentage.freq.peak.annotation

k4.ll.20.peak.annotation <- cbind(k4.ll.20[,1:3],data.frame(peak.annotation,target.genes))

head(k4.ll.20.peak.annotation)

write.table(x = k4.ll.20.peak.annotation,file = "tables/H3K4me3_LL_20C_peak_annotation.tsv",quote = F,sep = "\t",row.names = F)

all.k4.genes <- (unlist(sapply(X =k4.ll.20.peak.annotation$target.genes,
                                FUN = function(x) {strsplit(x,",")[[1]]})))
names(all.k4.genes) <- NULL
length(all.k4.genes)
length(all.k4.genes)/7668
```

We found 832 genes associated to H3K4me3 marks corresponding to 10.9% of the entire Ostreococcus genome. H3K4me3 is mostly located occupying the TSS (78.4%) with marginal percentages for internal regions of gene body (8.2%), intergenic regions (7%), entire gene bodies (5.3%) and TES (1.2%).  

Next we use metageneplots to determine the shape of the H3K27me3 over its occupied genes. We consider three different sets of marked genes, genes with their entire gene body marked, genes with H3K27 inside their gene body with no overlap in the TSS or TES and genes with H3K27me3 overlapping their TSS or TES

Next we generate the metageneplots for genes where the H3K27me3 occupies their entire gene body including TSS and TES. 

```{r k4_metageneplots_gene_body}
library(rtracklayer)
## Genes with their entire gene body marked
k4.gene.body <- (unlist(sapply(X = subset(k4.ll.20.peak.annotation, 
                                          peak.annotation == "gene.body")[["target.genes"]],
                                FUN = function(x) {strsplit(x,",")[[1]]})))
names(k4.gene.body) <- NULL
length(k4.gene.body)

## Genes with H3K27 inside their gene body with no overlap
## in the TSS or TES.
k4.internal.gene.body <- (unlist(sapply(X = subset(k4.ll.20.peak.annotation, 
                                                   peak.annotation == "internal.gene.body")[["target.genes"]],
                                         FUN = function(x) {strsplit(x,",")[[1]]})))
names(k4.internal.gene.body) <- NULL
length(k4.internal.gene.body)

## Genes with H3K27me3 overlapping their TSS
k4.tss <- (unlist(sapply(X = subset(k4.ll.20.peak.annotation, 
                                    peak.annotation == "TSS")[["target.genes"]],
                          FUN = function(x) {strsplit(x,",")[[1]]})))
names(k4.tss) <- NULL
length(k4.tss)

## Genes with H3K27me3 overlapping their TES
k4.tes <- (unlist(sapply(X = subset(k4.ll.20.peak.annotation, 
                                    peak.annotation == "TES")[["target.genes"]],
                          FUN = function(x) {strsplit(x,",")[[1]]})))
names(k4.tes) <- NULL
length(k4.tes)

metageneplot.data.k4.gene.body.1 <- 
 metageneplot(gene.names = k4.gene.body,
              genes = ostta.genes,
              bw.file.name = "chipseq_data/bw_data/h3k4me3/temp20/h3k4me3_20_chip_1.bw",
              ntile.txs = 10,
              ntile.gene.body = 20)
median.metageneplot.data.k4.gene.body.1 <- 
 apply(X = metageneplot.data.k4.gene.body.1,
       MARGIN = 2,
       FUN = median,na.rm = T)

metageneplot.data.k4.gene.body.2 <- 
 metageneplot(gene.names = k4.gene.body,
              genes = ostta.genes,
              bw.file.name = "chipseq_data/bw_data/h3k4me3/temp20/h3k4me3_20_chip_2.bw",
              ntile.txs = 10,
              ntile.gene.body = 20)
median.metageneplot.data.k4.gene.body.2 <- 
 apply(X = metageneplot.data.k4.gene.body.2,
       MARGIN = 2,
       FUN = median,na.rm = T)

metageneplot.data.k4.gene.body.3 <- 
 metageneplot(gene.names = k4.gene.body,
              genes = ostta.genes,
              bw.file.name = "chipseq_data/bw_data/h3k4me3/temp20/h3k4me3_20_chip_3.bw",
              ntile.txs = 10,
              ntile.gene.body = 20)
median.metageneplot.data.k4.gene.body.3 <- 
 apply(X = metageneplot.data.k4.gene.body.3,
       MARGIN = 2,
       FUN = median,na.rm = T)

median.metageneplot.data.k4.gene.body <- (median.metageneplot.data.k4.gene.body.1 + 
                                           median.metageneplot.data.k4.gene.body.2 + 
                                           median.metageneplot.data.k4.gene.body.3)/3

plot(smooth.spline(x=1:40,y=median.metageneplot.data.k4.gene.body,spar = 0.5),
     type="l",col="blue",lwd=3,ylim=c(0,70),axes=F,xlab="",ylab="")
axis(side = 1, at=c(0,10,30,40),labels = c("-2Kb","TSS","TES","2Kb"),las=2,lwd=2)
axis(side = 2,lwd=2)
```

Next we generate the metageneplots for genes where the H3K4me3 occupies part of the gene body without overlapping with the TSS and TES. 

```{r k4_metageneplots_internal_gene_body}
metageneplot.data.k4.internal.gene.body.1 <- 
 metageneplot(gene.names = k4.internal.gene.body,
              genes = ostta.genes,
              bw.file.name = "chipseq_data/bw_data/h3k4me3/temp20/h3k4me3_20_chip_1.bw",
              ntile.txs = 10,
              ntile.gene.body = 20)
median.metageneplot.data.k4.internal.gene.body.1 <- 
 apply(X = metageneplot.data.k4.internal.gene.body.1,
       MARGIN = 2,
       FUN = median,na.rm = T)


metageneplot.data.k4.internal.gene.body.2 <- 
 metageneplot(gene.names = k4.internal.gene.body,
              genes = ostta.genes,
              bw.file.name = "chipseq_data/bw_data/h3k4me3/temp20/h3k4me3_20_chip_2.bw",
              ntile.txs = 10,
              ntile.gene.body = 20)
median.metageneplot.data.k4.internal.gene.body.2 <- 
 apply(X = metageneplot.data.k4.internal.gene.body.2,
       MARGIN = 2,
       FUN = median,na.rm = T)


metageneplot.data.k4.internal.gene.body.3 <- 
 metageneplot(gene.names = k4.internal.gene.body,
              genes = ostta.genes,
              bw.file.name = "chipseq_data/bw_data/h3k4me3/temp20/h3k4me3_20_chip_3.bw",
              ntile.txs = 10,
              ntile.gene.body = 20)
median.metageneplot.data.k4.internal.gene.body.3 <- 
 apply(X = metageneplot.data.k4.internal.gene.body.3,
       MARGIN = 2,
       FUN = median,na.rm = T)

median.metageneplot.data.k4.internal.gene.body <- (median.metageneplot.data.k4.internal.gene.body.1 +
                                                    median.metageneplot.data.k4.internal.gene.body.2 +
                                                    median.metageneplot.data.k4.internal.gene.body.3)/3

plot(smooth.spline(x=1:40,y=median.metageneplot.data.k4.internal.gene.body,spar = 0.5),
     type="l",col="cyan",lwd=3,ylim=c(0,10),axes=F,xlab="",ylab="")
axis(side = 1, at=c(0,10,30,40),labels = c("-2Kb","TSS","TES","2Kb"),las=2,lwd=2)
axis(side = 2,lwd=2)
```

Next we generate the metageneplots for genes where the H3K4me3 occupies only the TSS and the beginning of the gene.  

```{r k4_metageneplots_tss}
metageneplot.data.k4.tss.1 <- 
 metageneplot(gene.names = k4.tss,
              genes = ostta.genes,
              bw.file.name = "chipseq_data/bw_data/h3k4me3/temp20/h3k4me3_20_chip_1.bw",
              ntile.txs = 10,
              ntile.gene.body = 20)
median.metageneplot.data.k4.tss.1 <- 
 apply(X = metageneplot.data.k4.tss.1,
       MARGIN = 2,
       FUN = median,na.rm = T)

metageneplot.data.k4.tss.2 <- 
 metageneplot(gene.names = k4.tss,
              genes = ostta.genes,
              bw.file.name = "chipseq_data/bw_data/h3k4me3/temp20/h3k4me3_20_chip_2.bw",
              ntile.txs = 10,
              ntile.gene.body = 20)
median.metageneplot.data.k4.tss.2 <- 
 apply(X = metageneplot.data.k4.tss.2,
       MARGIN = 2,
       FUN = median,na.rm = T)



metageneplot.data.k4.tss.3 <- 
 metageneplot(gene.names = k4.tss,
              genes = ostta.genes,
              bw.file.name = "chipseq_data/bw_data/h3k4me3/temp20/h3k4me3_20_chip_3.bw",
              ntile.txs = 10,
              ntile.gene.body = 20)
median.metageneplot.data.k4.tss.3 <- 
 apply(X = metageneplot.data.k4.tss.3,
       MARGIN = 2,
       FUN = median,na.rm = T)

median.metageneplot.data.k4.tss <- 
 (median.metageneplot.data.k4.tss.1 + 
   median.metageneplot.data.k4.tss.2 +
   median.metageneplot.data.k4.tss.3)/3

plot(smooth.spline(x=1:40,y=median.metageneplot.data.k4.tss,spar = 0.5),
     type="l",col="green",lwd=3,ylim=c(0,70),axes=F,xlab="",ylab="")
axis(side = 1, at=c(0,10,30,40),labels = c("-2Kb","TSS","TES","2Kb"),las=2,lwd=2)
axis(side = 2,lwd=2)
```

Next we generate the metageneplots for genes where the H3K4me3 occupies only the TES and the end of the gene.  

```{r k4_metageneplots_tes}
metageneplot.data.k4.tes.1 <- 
 metageneplot(gene.names = k4.tes,
              genes = ostta.genes,
              bw.file.name = "chipseq_data/bw_data/h3k4me3/temp20/h3k4me3_20_chip_1.bw",
              ntile.txs = 10,
              ntile.gene.body = 20)
median.metageneplot.data.k4.tes.1 <- 
 apply(X = metageneplot.data.k4.tes.1,
       MARGIN = 2,
       FUN = median,na.rm = T)



metageneplot.data.k4.tes.2 <- 
 metageneplot(gene.names = k4.tes,
              genes = ostta.genes,
              bw.file.name = "chipseq_data/bw_data/h3k4me3/temp20/h3k4me3_20_chip_2.bw",
              ntile.txs = 10,
              ntile.gene.body = 20)
median.metageneplot.data.k4.tes.2 <- 
 apply(X = metageneplot.data.k4.tes.2,
       MARGIN = 2,
       FUN = median,na.rm = T)


metageneplot.data.k4.tes.3 <- 
 metageneplot(gene.names = k4.tes,
              genes = ostta.genes,
              bw.file.name = "chipseq_data/bw_data/h3k4me3/temp20/h3k4me3_20_chip_3.bw",
              ntile.txs = 10,
              ntile.gene.body = 20)
median.metageneplot.data.k4.tes.3 <- 
 apply(X = metageneplot.data.k4.tes.3,
       MARGIN = 2,
       FUN = median,na.rm = T)


median.metageneplot.data.k4.tes <- 
 (median.metageneplot.data.k4.tes.1 + 
   median.metageneplot.data.k4.tes.2 +
   median.metageneplot.data.k4.tes.3)/3

plot(smooth.spline(x=1:40,y=median.metageneplot.data.k4.tes,spar = 0.5),
     type="l",col="red",lwd=3,ylim=c(0,20),axes=F,xlab="",ylab="")
axis(side = 1, at=c(0,10,30,40),labels = c("-2Kb","TSS","TES","2Kb"),las=2,lwd=2)
axis(side = 2,lwd=2)
```
In order to compare the different metageneplots are represented in the same plot.

```{r all_k4_metageneplots}
#jpeg(filename = "images/h3k4me3_metageneplots.jpg",res = 250,width = 900,height = 900)
plot(smooth.spline(x=1:40,y=median.metageneplot.data.k4.gene.body,spar = 0.5),
     type="l",col="blue",lwd=3,ylim=c(0,80),axes=F,xlab="",ylab="")
lines(smooth.spline(x=1:40,y=median.metageneplot.data.k4.internal.gene.body,spar = 0.5),
      type="l",col="cyan",lwd=3)
lines(smooth.spline(x=1:40,y=median.metageneplot.data.k4.tss,spar = 0.5),
      type="l",col="green",lwd=3)
lines(smooth.spline(x=1:40,y=median.metageneplot.data.k4.tes,spar = 0.5),
      type="l",col="red",lwd=3)
axis(side = 1, at=c(0,10,30,40),labels = c("-2Kb","TSS","TES","2Kb"),las=2,lwd=2)
axis(side = 2,lwd=2)
#dev.off()
```

It can be observed that the level of H3K4me3 mark varies depending on the part of gene that it occupies. Next we check if these differences are significant.  

```{r k4_metageneplots_significance}
max.metageneplot.data.k4.gene.body.1 <- 
 apply(X = metageneplot.data.k4.gene.body.1,
       MARGIN = 1,
       FUN = max,na.rm = T)
max.metageneplot.data.k4.gene.body.2 <- 
 apply(X = metageneplot.data.k4.gene.body.2,
       MARGIN = 1,
       FUN = max,na.rm = T)
max.metageneplot.data.k4.gene.body.3 <- 
 apply(X = metageneplot.data.k4.gene.body.3,
       MARGIN = 1,
       FUN = max,na.rm = T)

max.metageneplot.data.k4.gene.body <- (max.metageneplot.data.k4.gene.body.1 + 
                                   max.metageneplot.data.k4.gene.body.2 +
                                   max.metageneplot.data.k4.gene.body.3)/3

max.metageneplot.data.k4.internal.gene.body.1 <- 
 apply(X = metageneplot.data.k4.internal.gene.body.1,
       MARGIN = 1,
       FUN = max,na.rm = T)
max.metageneplot.data.k4.internal.gene.body.2 <- 
 apply(X = metageneplot.data.k4.internal.gene.body.2,
       MARGIN = 1,
       FUN = max,na.rm = T)
max.metageneplot.data.k4.internal.gene.body.3 <- 
 apply(X = metageneplot.data.k4.internal.gene.body.3,
       MARGIN = 1,
       FUN = max,na.rm = T)

max.metageneplot.data.k4.internal.gene.body <- (max.metageneplot.data.k4.internal.gene.body.1 + 
                                         max.metageneplot.data.k4.internal.gene.body.2 +
                                         max.metageneplot.data.k4.internal.gene.body.3)/3


max.metageneplot.data.k4.tss.1 <- 
 apply(X = metageneplot.data.k4.tss.1,
       MARGIN = 1,
       FUN = max,na.rm = T)
max.metageneplot.data.k4.tss.2 <- 
 apply(X = metageneplot.data.k4.tss.2,
       MARGIN = 1,
       FUN = max,na.rm = T)
max.metageneplot.data.k4.tss.3 <- 
 apply(X = metageneplot.data.k4.tss.3,
       MARGIN = 1,
       FUN = max,na.rm = T)
max.metageneplot.data.k4.tss <- (max.metageneplot.data.k4.tss.1 + 
                                   max.metageneplot.data.k4.tss.2 +
                                   max.metageneplot.data.k4.tss.3)/3

max.metageneplot.data.k4.tes.1 <- 
 apply(X = metageneplot.data.k4.tes.1,
       MARGIN = 1,
       FUN = max,na.rm = T)
max.metageneplot.data.k4.tes.2 <- 
 apply(X = metageneplot.data.k4.tes.2,
       MARGIN = 1,
       FUN = max,na.rm = T)
max.metageneplot.data.k4.tes.3 <- 
 apply(X = metageneplot.data.k4.tes.3,
       MARGIN = 1,
       FUN = max,na.rm = T)
max.metageneplot.data.k4.tes <- (max.metageneplot.data.k4.tes.1 + 
                                  max.metageneplot.data.k4.tes.2 +
                                  max.metageneplot.data.k4.tes.3)/3

shapiro.test(max.metageneplot.data.k4.gene.body)
shapiro.test(max.metageneplot.data.k4.internal.gene.body)
shapiro.test(max.metageneplot.data.k4.tss)
shapiro.test(max.metageneplot.data.k4.tes)

wilcox.test(x = max.metageneplot.data.k4.gene.body,
            y = max.metageneplot.data.k4.internal.gene.body,
            alternative = "greater")$p.value
wilcox.test(x = max.metageneplot.data.k4.gene.body,
            y = max.metageneplot.data.k4.tss,
            alternative = "greater")
wilcox.test(x = max.metageneplot.data.k4.gene.body,
            y = max.metageneplot.data.k4.tes,
            alternative = "greater")
wilcox.test(x = max.metageneplot.data.k4.tss,
            y = max.metageneplot.data.k4.tes)
```

Indeed the differences between all metageneplots are highly significant.

Now we analyse the gene expression level of H3K4me3 marked genes.

```{r h3k4me3_gene_expression}

non.k4.genes <- setdiff(names(gene.expression.20),c(k4.gene.body,k4.tss,k4.tes,k4.internal.gene.body))

summary(gene.expression.20[k4.gene.body])
summary(gene.expression.20[k4.internal.gene.body])
summary(gene.expression.20[k4.tss])
summary(gene.expression.20[k4.tes])
summary(gene.expression.20[non.k4.genes])

#jpeg(filename = "images/h3k24me3_expression_barplot_different_genes.jpg",res = 250,width = 600,height = 1000)
boxplot(gene.expression.20[k4.gene.body],
        gene.expression.20[k4.tss],
        gene.expression.20[k4.internal.gene.body],
        gene.expression.20[k4.tes],
        gene.expression.20[non.k4.genes],
        col=c("blue","green","cyan","red","grey"),
        outline=F)
#dev.off()

wilcox.test(x = gene.expression.20[k4.gene.body], 
            y = gene.expression.20[k4.tss],
            alternative = "greater")

wilcox.test(x = gene.expression.20[k4.tss], 
            y = gene.expression.20[k4.internal.gene.body],
            alternative = "greater")

wilcox.test(x = gene.expression.20[k4.internal.gene.body],
            y = gene.expression.20[k4.tes],
            alternative = "greater")

wilcox.test(x = gene.expression.20[k4.tes],
            y = gene.expression.20[non.k4.genes],
            alternative = "greater")
wilcox.test(x = gene.expression.20[k4.gene.body],
            y = gene.expression.20[non.k4.genes],
            alternative = "greater")
wilcox.test(x = gene.expression.20[k4.internal.gene.body],
            y = gene.expression.20[non.k4.genes],
            alternative = "greater")
wilcox.test(x = gene.expression.20[k4.tss],
            y = gene.expression.20[non.k4.genes],
            alternative = "greater")$p.value
wilcox.test(x = gene.expression.20[k4.tes],
            y = gene.expression.20[non.k4.genes],
            alternative = "greater")$p.value


```

We can see all H3K4me3 genes present levels of expression significantly higher than non-marked genes. Specifically, genes whose entire gene body or TSS is occupied by H3K4me3 are much higher expressed. Whereas, genes where only the TES or gene body is occupied by H3K4me3 are expressed at a significantly higher level than non-marked genes but with much lower significance level. 

In what follows we perform a functional analysis over the H3K4me3 marked genes. First we study the enrichment in GO terms.

```{r k4_go_enrichment}
library(clusterProfiler)
library(enrichplot)
library(org.Otauriv5.eg.db)
k4.enrichment <- enrichGO(gene = all.k4.genes, pvalueCutoff = 0.05,qvalueCutoff = 0.05,
                       OrgDb = org.Otauriv5.eg.db,
                       ont = "BP", 
                       keyType = "GID")
k4.enrichment.df <- as.data.frame(k4.enrichment)

write.table(x = k4.enrichment.df, file = "go_enrichment_k4.tsv",quote = F,sep = "\t",row.names = F)

treeplot(x = pairwise_termsim(k4.enrichment))

library(treemap)

revigo.names <- c("term_ID","description","frequency","value","uniqueness","dispensability","representative");
revigo.data <- rbind(c("GO:0006412","translation",4.38869169324396,41.635159951405136,0.6875841543305287,0,"translation"),
                     c("GO:0002181","cytoplasmic translation",0.3974763229665376,8.595948611417302,0.755633504893879,0.6792736,"translation"),
                     c("GO:0009059","macromolecule biosynthetic process",16.15219652221219,32.13388504723149,0.8406622568296607,0.37915401,"translation"),
                     c("GO:0043603","amide metabolic process",6.707376287050344,36.45011550587449,0.9109747285969385,0.18706276,"translation"),
                     c("GO:0009765","photosynthesis, light harvesting",0.02241434707504848,4.839062538560749,0.952328578586525,0.07239153,"photosynthesis, light harvesting"),
                     c("GO:0015979","photosynthesis",0.228607115192195,7.075542911759376,0.941812778414762,0.0906509,"photosynthesis"),
                     c("GO:0048518","positive regulation of biological process",4.759580864033397,2.926034087880324,0.8707557167557012,-0,"positive regulation of biological process"),
                     c("GO:0010557","positive regulation of macromolecule biosynthetic process",2.093180586483937,2.989915809683905,0.7484373322239899,0.27791144,"positive regulation of biological process"),
                     c("GO:0031323","regulation of cellular metabolic process",14.819661863282441,3.0711442275048717,0.7768013068864807,0.54794604,"positive regulation of biological process"),
                     c("GO:0140694","non-membrane-bounded organelle assembly",0.7232459602662196,6.425049931933804,0.6134007334542475,0.01308302,"ribosome biogenesis"),
                     c("GO:0000027","ribosomal large subunit assembly",0.012821578346646382,3.275341540820152,0.6307593041874131,0.61078702,"ribosome biogenesis"),
                     c("GO:0022613","ribonucleoprotein complex biogenesis",2.5375929564430133,4.856610166214762,0.6332784349267674,0.65931649,"ribosome biogenesis"),
                     c("GO:0042254","ribosome biogenesis",2.136224808753416,5.364416283328431,0.566141578351784,0.5650077,"ribosome biogenesis"),
                     c("GO:0070925","organelle assembly",1.3561320812847646,3.9914667506842227,0.6213427911227302,0.65542123,"ribosome biogenesis"),
                     c("GO:0071826","protein-RNA complex organization",0.6518695497816881,3.3975782710402287,0.7204190675227414,0.36838395,"ribosome biogenesis"));

stuff <- data.frame(revigo.data);
names(stuff) <- revigo.names;

stuff$value <- as.numeric( as.character(stuff$value) );
stuff$frequency <- as.numeric( as.character(stuff$frequency) );
stuff$uniqueness <- as.numeric( as.character(stuff$uniqueness) );
stuff$dispensability <- as.numeric( as.character(stuff$dispensability) );

#jpeg(filename = "images/h3k4me3_revigo_treemap.jpg",res = 300,width = 1400,height = 900)

treemap(
 stuff,
 index = c("representative","description"),
 vSize = "value",
 type = "categorical",
 vColor = "representative",
 title = "",
 inflate.labels = TRUE,      
 lowerbound.cex.labels = 0,   
 position.legend = "none"
)

#dev.off()
```



## H3K27me3 and H3K4me3 common marked genes

The set of H3K27me3 and H3K4me3 genes only overlapped marginally. 

```{r venn_diagram_k27_k4}
overlap.k4.k27 <- read.table(file="chipseq_data/bed_data/h3k27me3_h3k4me3_ll_20.bed",header=F,sep = "\t",as.is = T)

max(overlap.k4.k27$V3 - overlap.k4.k27$V2)
summary(overlap.k4.k27$V3 - overlap.k4.k27$V2)
nrow(overlap.k4.k27)

library(VennDiagram)

#jpeg(filename = "images/h3k27me3_h3k4me3_venndiagram.jpg",res = 250,width = 1500,height = 1500)
grid.newpage()
draw.pairwise.venn(area1 = length(all.k27.genes),cat.pos = c(180,180),cat.dist = 0.05,
                   area2 = length(all.k4.genes),
                   cross.area = length(intersect(all.k27.genes,all.k4.genes)),
                   lwd = 3,category = c("H3K27me3","H3K4me3"),euler.d = T,
                   col = c("blue","darkorange"),
                   fill = c(" blue","darkorange"),alpha = 0.5,
                   cex = 2,
                   cat.cex = 2)
#dev.off()
```

Only 19 genomic regions were found simultaneously marked by H3K27me3 and H3K4me3. The length of the overlap was never longer than 400bp and on average was less than 200bp. Taking into account that the length of DNA wrapped around a single nucleosome is approximately 150bp, no more than one or two nucleosomes appeared marked simultaneously by H3K27me3 and H3K4me3. 

Only 37 genes were identified as bivalent simultaneously marked with H3K27me3 and H3K4me3. Next we explore the positioning of these two marks in these genes and their level of expresssion.

```{r bivalent_marks_positioning}
k4.k27.genes <- intersect(all.k4.genes, all.k27.genes)
k4.k27.gene.annotation <- subset(ll.20.peak.annotation, target.genes %in% k4.k27.genes)
100*table(k4.k27.gene.annotation[!duplicated(k4.k27.gene.annotation$target.genes),][["peak.annotation"]])/nrow(k4.k27.gene.annotation)
```

In most bivalent genes H3K27me3 is located in inner regions of gene bodies (58.5%). In 14.6% of the bivalent genes H3K27me3 is located only at the TES. Whereas in only 9.8% and 4.9% H3K27me3 occupies the entire gene body or only the TSS, respectively. 


```{r bivalent_gene_expression}
non.k4.k27.genes <- intersect(non.k4.genes,non.k27.genes)
summary(gene.expression.20[all.k4.genes])
summary(gene.expression.20[all.k27.genes])
summary(gene.expression.20[k4.k27.genes])
summary(gene.expression.20[non.k4.k27.genes])

#jpeg(filename = "images/bivalent_expression_barplot.jpg",res = 250,width = 600,height = 1000)
boxplot(gene.expression.20[all.k27.genes],
        gene.expression.20[non.k4.k27.genes],
        gene.expression.20[k4.k27.genes],
        gene.expression.20[all.k4.genes],
        col=c("blue","grey",colorRampPalette(c('blue','darkorange'))(3)[2],"darkorange"),outline=F)
#dev.off()

wilcox.test(x = gene.expression.20[k4.k27.genes], y = gene.expression.20[non.k4.k27.genes])
wilcox.test(x = gene.expression.20[all.k27.genes], y = gene.expression.20[k4.k27.genes],alternative = "less")
wilcox.test(x = gene.expression.20[k4.k27.genes], y = gene.expression.20[all.k4.genes],alternative = "less")
```


## H3K27me3 response to diel cycles and constant light

```{r load_data_initial_exploration}
mapped.reads.k27.20c.1 <- 3146662/1E6 
mapped.reads.k27.20c.2 <- 5697849/1E6
mapped.reads.k27.20c.3 <- 5157582/1E6
mapped.reads.k27.20c.zt8.1 <- 2254676/1E6 
mapped.reads.k27.20c.zt8.2 <- 2639017/1E6
mapped.reads.k27.20c.zt16.1 <- 3605883/1E6
mapped.reads.k27.20c.zt16.2 <- 3405826/1E6


read.counts.k27.temp20c.1 <- read.table(file ="chipseq_data/consensus_counts/h3k27me3_ll_ld_consensus_ll1_counts.bed", header = F, sep = "\t" )
read.counts.k27.temp20c.2 <- read.table(file = "chipseq_data/consensus_counts/h3k27me3_ll_ld_consensus_ll2_counts.bed", header = F, sep = "\t" )
read.counts.k27.temp20c.3 <- read.table(file = "chipseq_data/consensus_counts/h3k27me3_ll_ld_consensus_ll3_counts.bed", header = F, sep = "\t" )

read.counts.k27.20c.zt8.1 <- read.table(file = "chipseq_data/consensus_counts/h3k27me3_ll_ld_consensus_zt8_1_counts.bed", header = F, sep = "\t" )
read.counts.k27.20c.zt8.2 <- read.table(file = "chipseq_data/consensus_counts/h3k27me3_ll_ld_consensus_zt8_2_counts.bed", header = F, sep = "\t" )

read.counts.k27.20c.zt16.1 <- read.table(file = "chipseq_data/consensus_counts/h3k27me3_ll_ld_consensus_zt16_1_counts.bed", header = F, sep = "\t" )
read.counts.k27.20c.zt16.2 <- read.table(file = "chipseq_data/consensus_counts/h3k27me3_ll_ld_consensus_zt16_2_counts.bed", header = F, sep = "\t" )



```

## H3K27me3 response to temperature variation

```{r load_data}

mapped.reads.k27.10c.1 <- 6764067/1E6 
mapped.reads.k27.10c.2 <- 6429853/1E6
mapped.reads.k27.10c.3 <- 6625634/1E6

mapped.reads.k27.14c.1 <- 24537863/1E6 
mapped.reads.k27.14c.2 <- 6592650/1E6
mapped.reads.k27.14c.3 <- 4317544/1E6

mapped.reads.k27.20c.1 <- 3146662/1E6 
mapped.reads.k27.20c.2 <- 7702115/1E6
mapped.reads.k27.20c.3 <- 7102648/1E6

mapped.reads.k27.26c.1 <- 5635098/1E6 
mapped.reads.k27.26c.2 <- 4986577/1E6
mapped.reads.k27.26c.3 <-4586942/1E6

read.counts.k27.temp10c.1 <- read.table(file = "chipseq_data/bed_data/h3k27me3/temp10c/h3k27me3_10_chip_1_counts.bed", header = F, sep = "\t" )
read.counts.k27.temp10c.2 <- read.table(file = "chipseq_data/bed_data/h3k27me3/temp10c/h3k27me3_10_chip_2_counts.bed", header = F, sep = "\t" )
read.counts.k27.temp10c.3 <- read.table(file = "chipseq_data/bed_data/h3k27me3/temp10c/h3k27me3_10_chip_3_counts.bed", header = F, sep = "\t" )

read.counts.k27.temp14c.1 <- read.table(file = "chipseq_data/bed_data/h3k27me3/temp14c/h3k27me3_14_chip_1_counts.bed", header = F, sep = "\t" )
read.counts.k27.temp14c.2 <- read.table(file = "chipseq_data/bed_data/h3k27me3/temp14c/h3k27me3_14_chip_2_counts.bed", header = F, sep = "\t" )
read.counts.k27.temp14c.3 <- read.table(file = "chipseq_data/bed_data/h3k27me3/temp14c/h3k27me3_14_chip_3n_counts.bed", header = F, sep = "\t" )

read.counts.k27.temp20c.1 <- read.table(file = "chipseq_data/bed_data/h3k27me3/temp20c/h3k27me3_20_chip_1_counts.bed", header = F, sep = "\t" )
read.counts.k27.temp20c.2 <- read.table(file = "chipseq_data/bed_data/h3k27me3/temp20c/h3k27me3_20_chip_2n_counts.bed", header = F, sep = "\t" )
read.counts.k27.temp20c.3 <- read.table(file = "chipseq_data/bed_data/h3k27me3/temp20c/h3k27me3_20_chip_3n_counts.bed", header = F, sep = "\t" )

read.counts.k27.temp26c.1 <- read.table(file = "chipseq_data/bed_data/h3k27me3/temp26c/h3k27me3_26_chip_1_counts.bed", header = F, sep = "\t" )
read.counts.k27.temp26c.2 <- read.table(file = "chipseq_data/bed_data/h3k27me3/temp26c/h3k27me3_26_chip_2_counts.bed", header = F, sep = "\t" )
read.counts.k27.temp26c.3 <- read.table(file = "chipseq_data/bed_data/h3k27me3/temp26c/h3k27me3_26_chip_3_counts.bed", header = F, sep = "\t" )


```



```{r}
peak.cpm <- matrix(data = NA, nrow = nrow(read.counts.k27.temp10c.1), ncol = 15 )
colnames(peak.cpm) <-c("chr","start","end","k27_temp10_1","k27_temp10_2","k27_temp10_3", "k27_temp14_1","k27_temp14_2","k27_temp14_3","k27_temp20_1","k27_temp20_2","k27_temp20_3","k27_temp26_1","k27_temp26_2","k27_temp26_3")

peak.cpm[,1]<- read.counts.k27.temp10c.1[,1]
peak.cpm[,2]<- read.counts.k27.temp10c.1[,2]
peak.cpm[,3]<- read.counts.k27.temp10c.1[,3]

peak.cpm[,4]<- read.counts.k27.temp10c.1[,4]/mapped.reads.k27.10c.1
peak.cpm[,5]<- read.counts.k27.temp10c.2[,4]/mapped.reads.k27.10c.2
peak.cpm[,6]<- read.counts.k27.temp10c.3[,4]/mapped.reads.k27.10c.3

peak.cpm[,7]<- read.counts.k27.temp14c.1[,4]/mapped.reads.k27.14c.1
peak.cpm[,8]<- read.counts.k27.temp14c.2[,4]/mapped.reads.k27.14c.2
peak.cpm[,9]<- read.counts.k27.temp14c.3[,4]/mapped.reads.k27.14c.3

peak.cpm[,10]<- read.counts.k27.temp20c.1[,4]/mapped.reads.k27.20c.1
peak.cpm[,11]<- read.counts.k27.temp20c.2[,4]/mapped.reads.k27.20c.2
peak.cpm[,12]<- read.counts.k27.temp20c.3[,4]/mapped.reads.k27.20c.3

peak.cpm[,13]<- read.counts.k27.temp26c.1[,4]/mapped.reads.k27.26c.1
peak.cpm[,14]<- read.counts.k27.temp26c.2[,4]/mapped.reads.k27.26c.2
peak.cpm[,15]<- read.counts.k27.temp26c.3[,4]/mapped.reads.k27.26c.3

head(peak.cpm)
dim(peak.cpm)


peak.cpm <- as.data.frame(peak.cpm)
```

#Scatterplots representing changes observed between temperature conditions

```{r}
plot(x = log2((peak.cpm$k27_temp10_1 + peak.cpm$k27_temp10_2 +peak.cpm$k27_temp10_3)/3), y= log2((peak.cpm$k27_temp14_1 + peak.cpm$k27_temp14_2 + peak.cpm$k27_temp14_3)/3), pch = 20,xlab="H3K27me3 10ºC (CPM)",ylab="H3K27me3 14ºC (CPM)",main="14ºC vs 10ºC")

plot(x = log2((peak.cpm$k27_temp10_1 + peak.cpm$k27_temp10_2 +peak.cpm$k27_temp10_3)/3), y= log2((peak.cpm$k27_temp20_1 + peak.cpm$k27_temp20_2 + peak.cpm$k27_temp20_3)/3), pch = 20,xlab="H3K27me3 10ºC (CPM)",ylab="H3K27me3 20ºC (CPM)",main="20ºC vs 10ºC")
lines(x = c(0,20), y = c(0,20), type = "l", col="blue")

plot(x = log2((peak.cpm$k27_temp10_1 + peak.cpm$k27_temp10_2 +peak.cpm$k27_temp10_3)/3), y= log2((peak.cpm$k27_temp26_1 + peak.cpm$k27_temp26_2 + peak.cpm$k27_temp26_3)/3), pch = 20,xlab="H3K27me3 10ºC (CPM)",ylab="H3K27me3 26ºC (CPM)",main="26ºC vs 10ºC")
lines(x = c(0,20), y = c(0,20), type = "l", col="dodgerblue2"
      
```

```{r}

wilcox.test(x = (peak.cpm$k27_temp10_1 + peak.cpm$k27_temp10_2 +peak.cpm$k27_temp10_3)/3, y = (peak.cpm$k27_temp14_1 + peak.cpm$k27_temp14_2 +peak.cpm$k27_temp14_3)/3,alternative = "less")

wilcox.test(x = (peak.cpm$k27_temp10_1 + peak.cpm$k27_temp10_2 +peak.cpm$k27_temp10_3)/3, y = (peak.cpm$k27_temp20_1 + peak.cpm$k27_temp20_2 +peak.cpm$k27_temp20_3)/3,alternative = "less")

wilcox.test(x = (peak.cpm$k27_temp10_1 + peak.cpm$k27_temp10_2 +peak.cpm$k27_temp10_3)/3, y = (peak.cpm$k27_temp26_1 + peak.cpm$k27_temp26_2 +peak.cpm$k27_temp26_3)/3,alternative = "less")

wilcox.test(x = (peak.cpm$k27_temp14_1 + peak.cpm$k27_temp14_2 +peak.cpm$k27_temp14_3)/3, y = (peak.cpm$k27_temp20_1 + peak.cpm$k27_temp20_2 +peak.cpm$k27_temp20_3)/3,alternative = "less")

wilcox.test(x = (peak.cpm$k27_temp20_1 + peak.cpm$k27_temp20_2 +peak.cpm$k27_temp20_3)/3, y = (peak.cpm$k27_temp26_1 + peak.cpm$k27_temp26_2 +peak.cpm$k27_temp26_3)/3,alternative = "less")

boxplot((peak.cpm$k27_temp10_1 + peak.cpm$k27_temp10_2 +peak.cpm$k27_temp10_3)/3,
        (peak.cpm$k27_temp14_1 + peak.cpm$k27_temp14_2 +peak.cpm$k27_temp14_3)/3,
        (peak.cpm$k27_temp20_1 + peak.cpm$k27_temp20_2 +peak.cpm$k27_temp20_3)/3,
        (peak.cpm$k27_temp26_1 + peak.cpm$k27_temp26_2 + peak.cpm$k27_temp26_3)/3, y= c(0,5000),  outline=F,col=c("#00FFFF","#6699FF", "#CC33FF","#FF0000", "grey"),ylab="CPM",names=c("10ºC","14ºC","20ºC","26ºC"),las=2,cex.lab=1, cex.axis= 0.8)

points(3,1290,pch=8,cex=0.7)
points(3.9,1980,pch=8,cex=0.7)
points(4,1980,pch=8,cex=0.7)
points(4.1,1980,pch=8,cex=0.7)
```
```{r}
peak.names <- paste(paste(peak.cpm$chr, peak.cpm$start, sep = "_"), peak.cpm$end, sep = "_")
log2.peak.cpm <- matrix(data = NA, nrow = nrow(read.counts.k27.temp10c.1), ncol = 12)
colnames(log2.peak.cpm) <- c("k27_temp10_1","k27_temp10_2","k27_temp10_3", "k27_temp14_1","k27_temp14_2","k27_temp14_3","k27_temp20_1","k27_temp20_2","k27_temp20_3","k27_temp26_1","k27_temp26_2","k27_temp26_3")
rownames(log2.peak.cpm) <- peak.names 

log2.peak.cpm[,"k27_temp10_1"] <- log2(peak.cpm$k27_temp10_1)
log2.peak.cpm[,"k27_temp10_2"] <- log2(peak.cpm$k27_temp10_2)
log2.peak.cpm[,"k27_temp10_3"] <- log2(peak.cpm$k27_temp10_3)

log2.peak.cpm[,"k27_temp14_1"] <- log2(peak.cpm$k27_temp14_1)
log2.peak.cpm[,"k27_temp14_2"] <- log2(peak.cpm$k27_temp14_2)
log2.peak.cpm[,"k27_temp14_3"] <- log2(peak.cpm$k27_temp14_3)

log2.peak.cpm[,"k27_temp20_1"] <- log2(peak.cpm$k27_temp20_1)
log2.peak.cpm[,"k27_temp20_2"] <- log2(peak.cpm$k27_temp20_2)
log2.peak.cpm[,"k27_temp20_3"] <- log2(peak.cpm$k27_temp20_3)

log2.peak.cpm[,"k27_temp26_1"] <- log2(peak.cpm$k27_temp26_1)
log2.peak.cpm[,"k27_temp26_2"] <- log2(peak.cpm$k27_temp26_2)
log2.peak.cpm[,"k27_temp26_3"] <- log2(peak.cpm$k27_temp26_3)


head(log2.peak.cpm)

```


#Differential occupancy analysis
```{r limma}
library(limma)

experimental.design <- model.matrix(~ -1+factor(c(1,1,1,2,2,2,3,3,3,4,4,4)))
colnames(experimental.design) <- c("k27_temp10",
                                   "k27_temp14",
                                   "k27_temp20",
                                   "k27_temp26")

experimental.design

linear.fit <- lmFit(log2.peak.cpm, experimental.design)

head(log2.peak.cpm)

contrast.matrix <- makeContrasts(k27_temp14-k27_temp10,
                                 k27_temp20-k27_temp10,
                                 k27_temp26-k27_temp10,
                                 levels=c("k27_temp10",
                                          "k27_temp14",
                                          "k27_temp20",
                                          "k27_temp26"))

contrast.linear.fit <- contrasts.fit(linear.fit, contrast.matrix)
contrast.results <- eBayes(contrast.linear.fit)
```


```{r temp14c-temp10c contrast}
temp14c.vs.temp10c <- topTable(contrast.results, number=nrow(log2.peak.cpm),coef=1)#,sort.by="logFC")
head(temp14c.vs.temp10c)
peak.ids.14c.10c <- rownames(temp14c.vs.temp10c) 
fold.change.14c.10c <- temp14c.vs.temp10c$logFC
q.value.14c.10c <- temp14c.vs.temp10c$adj.P.Val

activated.peaks.temp14c.vs.temp10c <- peak.ids.14c.10c[fold.change.14c.10c > 1 & q.value.14c.10c < 0.05]
repressed.peaks.temp14c.vs.temp10c <- peak.ids.14c.10c[fold.change.14c.10c < - 1 & q.value.14c.10c < 0.05]

length(activated.peaks.temp14c.vs.temp10c)
length(repressed.peaks.temp14c.vs.temp10c)

length(activated.peaks.temp14c.vs.temp10c)/length(peak.names)

```

```{r temp20c-temp10c contrast}
temp20c.vs.temp10c <- topTable(contrast.results, number=nrow(log2.peak.cpm),coef=2,sort.by="logFC")
head(temp20c.vs.temp10c)
peak.ids.20c.10c <- rownames(temp20c.vs.temp10c)
fold.change.20c.10c <- temp20c.vs.temp10c$logFC
q.value.20c.10c <- temp20c.vs.temp10c$adj.P.Val

activated.peaks.temp20c.vs.temp10c <- peak.ids.20c.10c[fold.change.20c.10c > 1 & q.value.20c.10c < 0.05]
repressed.peaks.temp20c.vs.temp10c <- peak.ids.20c.10c[fold.change.20c.10c < - 1 & q.value.20c.10c < 0.05]

length(activated.peaks.temp20c.vs.temp10c)
length(repressed.peaks.temp20c.vs.temp10c)

length(activated.peaks.temp20c.vs.temp10c)/length(peak.names)
```


```{r temp26c-temp10c contrast}
temp26c.vs.temp10c <- topTable(contrast.results, number=nrow(log2.peak.cpm),coef=3,sort.by="logFC")
head(temp26c.vs.temp10c)
peak.ids.26c.10c <- rownames(temp26c.vs.temp10c) 
fold.change.26c.10c <- temp26c.vs.temp10c$logFC
q.value.26c.10c <- temp26c.vs.temp10c$adj.P.Val

activated.peaks.temp26c.vs.temp10c <- peak.ids.26c.10c[fold.change.26c.10c > 1 & q.value.26c.10c < 0.05]
repressed.peaks.temp26c.vs.temp10c <- peak.ids.26c.10c[fold.change.26c.10c < - 1 & q.value.26c.10c < 0.05]

length(activated.peaks.temp26c.vs.temp10c)
length(repressed.peaks.temp26c.vs.temp10c)

length(activated.peaks.temp26c.vs.temp10c)/length(peak.names)

library(TxDb.Otauri.JGI)
txdb <- TxDb.Otauri.JGI
ostta.genes <- as.data.frame(genes(txdb))
head(ostta.genes)


matrix.activated.peaks.temp26c.vs.temp10c <- as.matrix(t(as.data.frame(strsplit(x = activated.peaks.temp26c.vs.temp10c, split = "_"))))
rownames(matrix.activated.peaks.temp26c.vs.temp10c) <- NULL
head(matrix.activated.peaks.temp26c.vs.temp10c)
write.table(matrix.activated.peaks.temp26c.vs.temp10c, file = "chipseq_results/h3k27me3/limma.k27.activated.peaks.temp26.vs.temp10.bed", sep = "\t", quote = FALSE, row.names = FALSE, col.names = FALSE)

activated.peaks.26vs10 <- read.table(file = "chipseq_results/h3k27me3/limma.k27.activated.peaks.temp26.vs.temp10.bed", sep = "\t")
colnames(activated.peaks.26vs10) <- c("chr", "start", "end")

activated.genes.26.10 <- vector(mode = "character",length = nrow(activated.peaks.26vs10))
peak.annotation.26.10 <- vector(mode = "character",length = nrow(activated.peaks.26vs10))

for(i in 1:nrow(activated.peaks.26vs10))
{
 current.chr <- activated.peaks.26vs10[i,1]
 current.start <- activated.peaks.26vs10[i,2]
 current.end <- activated.peaks.26vs10[i,3]

 complete.gene.body <- subset(ostta.genes, 
                              seqnames == current.chr & 
                               start >= current.start & 
                               end <= current.end)
 
 overlap.tss.pos.strand <- subset(ostta.genes, 
                              seqnames == current.chr & 
                              start >= current.start & start <= current.end & 
                              strand == "+")
 
 overlap.tss.neg.strand <- subset(ostta.genes, 
                              seqnames == current.chr & 
                              end >= current.start & end <= current.end & 
                              strand == "-")
 
 overlap.tes.pos.strand <- subset(ostta.genes, 
                              seqnames == current.chr & 
                              end >= current.start & end <= current.end & 
                              strand == "+")
 
 overlap.tes.neg.strand <- subset(ostta.genes, 
                              seqnames == current.chr & 
                              start >= current.start & start <= current.end & 
                              strand == "-")

 inside.gene.body <- subset(ostta.genes, seqnames == current.chr & 
                             start <= current.start & end >= current.end)
 
 
 if(nrow(complete.gene.body) > 0)
 {
  activated.genes.26.10[i] <- paste(unlist(complete.gene.body$gene_id),collapse = ",")
  peak.annotation.26.10[i] <- "gene.body"
 } else if(nrow(overlap.tss.pos.strand) > 0)
 {
  activated.genes.26.10[i] <- paste(unlist(overlap.tss.pos.strand$gene_id),collapse = ",")
  peak.annotation.26.10[i] <- "TSS"
 } else if(nrow(overlap.tss.neg.strand) > 0)
 {
  activated.genes.26.10[i] <- paste(unlist(overlap.tss.neg.strand$gene_id),collapse = ",")
  peak.annotation.26.10[i] <- "TSS"
 } else if(nrow(overlap.tes.pos.strand) > 0)
 {
  activated.genes.26.10[i] <- paste(unlist(overlap.tes.pos.strand$gene_id),collapse = ",")
  peak.annotation.26.10[i] <- "TES"
 } else if(nrow(overlap.tes.neg.strand) > 0)
 {
  activated.genes.26.10[i] <- paste(unlist(overlap.tes.neg.strand$gene_id),collapse = ",")
  peak.annotation.26.10[i] <- "TES"
 } else if(nrow(inside.gene.body) > 0)
 {
  activated.genes.26.10[i] <- paste(unlist(inside.gene.body$gene_id),collapse = ",")
  peak.annotation.26.10[i] <- "internal.gene.body"
 } else
 {
  peak.annotation.26.10[i] <- "intergenic"
 }
}


temp26vs10.peak.annotation <- cbind(activated.peaks.26vs10,data.frame(peak.annotation.26.10,activated.genes.26.10))
head(temp26vs10.peak.annotation)


write.table(x = temp26vs10.peak.annotation,file = "tables/H3K27me3_26vs10_peak_annotation.tsv",quote = F,sep = "\t",row.names = F) 
all.k27.26vs10.genes <- (unlist(sapply(X =temp26vs10.peak.annotation$activated.genes.26.10,
FUN = function(x) {strsplit(x,",")[[1]]})))
names(all.k27.26vs10.genes) <- NULL
length(all.k27.26vs10.genes)

length(all.k27.26vs10.genes)/7668

```

## H3K4me3 response to temperature variation

```{r load_data}
mapped.reads.k4.10c.1 <- 4660652/1E6 
mapped.reads.k4.10c.2 <- 4572532/1E6
mapped.reads.k4.10c.3 <- 4490999/1E6

mapped.reads.k4.14c.1 <- 4609469/1E6 
mapped.reads.k4.14c.2 <- 5034594/1E6
mapped.reads.k4.14c.3 <- 5174666/1E6

mapped.reads.k4.20c.1 <- 6832850/1E6 
mapped.reads.k4.20c.2 <- 5993486/1E6
mapped.reads.k4.20c.3 <- 6701422/1E6

mapped.reads.k4.26c.1 <- 6059862/1E6 
mapped.reads.k4.26c.2 <- 5263311/1E6
mapped.reads.k4.26c.3 <- 4251439/1E6


# This was marked as a conflict during merging. it needs to be checked (start)
read.counts.k4.temp10c.1 <- read.table(file = "chipseq_data/bed_data/h3k4me3/temp10c/h3k4me3_10_chip_1_counts.bed", header = F, sep = "\t" )
read.counts.k4.temp10c.2 <- read.table(file = "chipseq_data/bed_data/h3k4me3/temp10c/h3k4me3_10_chip_2_counts.bed", header = F, sep = "\t" )
read.counts.k4.temp10c.3 <- read.table(file = "chipseq_data/bed_data/h3k4me3/temp10c/h3k4me3_10_chip_3_counts.bed", header = F, sep = "\t" )

read.counts.k4.temp14c.1 <- read.table(file = "chipseq_data/bed_data/h3k4me3/temp14c/h3k4me3_14_chip_1_counts.bed", header = F, sep = "\t" )
read.counts.k4.temp14c.2 <- read.table(file = "chipseq_data/bed_data/h3k4me3/temp14c/h3k4me3_14_chip_2_counts.bed", header = F, sep = "\t" )
read.counts.k4.temp14c.3 <- read.table(file = "chipseq_data/bed_data/h3k4me3/temp14c/h3k4me3_14_chip_3_counts.bed", header = F, sep = "\t" )
# This was marked as a conflict during merging. it needs to be checked (end)

read.counts.k4.temp20c.1 <- read.table(file = "chipseq_data/bed_data/h3k4me3/temp20c/h3k4me3_20_chip_1_counts.bed", header = F, sep = "\t" )
read.counts.k4.temp20c.2 <- read.table(file = "chipseq_data/bed_data/h3k4me3/temp20c/h3k4me3_20_chip_2_counts.bed", header = F, sep = "\t" )
read.counts.k4.temp20c.3 <- read.table(file = "chipseq_data/bed_data/h3k4me3/temp20c/h3k4me3_20_chip_3_counts.bed", header = F, sep = "\t" )

read.counts.k4.temp26c.1 <- read.table(file = "chipseq_data/bed_data/h3k4me3/temp26c/h3k4me3_26_chip_1_counts.bed", header = F, sep = "\t" )
read.counts.k4.temp26c.2 <- read.table(file = "chipseq_data/bed_data/h3k4me3/temp26c/h3k4me3_26_chip_2_counts.bed", header = F, sep = "\t" )
read.counts.k4.temp26c.3 <- read.table(file = "chipseq_data/bed_data/h3k4me3/temp26c/h3k4me3_26_chip_3_counts.bed", header = F, sep = "\t" )

```

```{r}
peak.cpm.k4 <- matrix(data = NA, nrow = nrow(read.counts.k4.temp10c.1), ncol = 15)
colnames(peak.cpm.k4) <-c("chr","start","end","k4_temp10_1","k4_temp10_2","k4_temp10_3", "k4_temp14_1","k4_temp14_2","k4_temp14_3","k4_temp20_1","k4_temp20_2","k4_temp20_3","k4_temp26_1","k4_temp26_2","k4_temp26_3")

peak.cpm.k4[,1]<- read.counts.k4.temp10c.1[,1]
peak.cpm.k4[,2]<- read.counts.k4.temp10c.1[,2]
peak.cpm.k4[,3]<- read.counts.k4.temp10c.1[,3]

peak.cpm.k4[,4]<- read.counts.k4.temp10c.1[,4]/mapped.reads.k4.10c.1
peak.cpm.k4[,5]<- read.counts.k4.temp10c.2[,4]/mapped.reads.k4.10c.2
peak.cpm.k4[,6]<- read.counts.k4.temp10c.3[,4]/mapped.reads.k4.10c.3

peak.cpm.k4[,7]<- read.counts.k4.temp14c.1[,4]/mapped.reads.k4.14c.1
peak.cpm.k4[,8]<- read.counts.k4.temp14c.2[,4]/mapped.reads.k4.14c.2
peak.cpm.k4[,9]<- read.counts.k4.temp14c.3[,4]/mapped.reads.k4.14c.3

peak.cpm.k4[,10]<- read.counts.k4.temp20c.1[,4]/mapped.reads.k4.20c.1
peak.cpm.k4[,11]<- read.counts.k4.temp20c.2[,4]/mapped.reads.k4.20c.2
peak.cpm.k4[,12]<- read.counts.k4.temp20c.3[,4]/mapped.reads.k4.20c.3

peak.cpm.k4[,13]<- read.counts.k4.temp26c.1[,4]/mapped.reads.k4.26c.1
peak.cpm.k4[,14]<- read.counts.k4.temp26c.2[,4]/mapped.reads.k4.26c.2
peak.cpm.k4[,15]<- read.counts.k4.temp26c.3[,4]/mapped.reads.k4.26c.3

head(peak.cpm.k4)

peak.cpm.k4 <- as.data.frame(peak.cpm.k4)

```


#Scatterplots representing changes observed between temperature conditions

```{r}
plot(x = log2((peak.cpm.k4$k4_temp10_1 + peak.cpm.k4$k4_temp10_2 +peak.cpm.k4$k4_temp10_3)/3), y= log2((peak.cpm.k4$k4_temp14_1 + peak.cpm.k4$k4_temp14_2 + peak.cpm.k4$k4_temp14_3)/3), pch = 20,xlab="H3K4me3 10ºC (CPM)",ylab="H3K4me3 14ºC (CPM)",main="14ºC vs 10ºC")

plot(x = log2((peak.cpm.k4$k4_temp10_1 + peak.cpm.k4$k4_temp10_1 +peak.cpm.k4$k4_temp10_1)/3), y= log2((peak.cpm.k4$k4_temp20_1 + peak.cpm.k4$k4_temp20_2 + peak.cpm.k4$k4_temp20_3)/3), pch = 20,xlab="H3K4me3 10ºC (CPM)",ylab="H3K4me3 20ºC (CPM)",main="20ºC vs 10ºC")
lines(x = c(0,20), y = c(0,20), type = "l", col="blue")

plot(x = log2((peak.cpm.k4$k4_temp10_1 + peak.cpm.k4$k4_temp10_1 +peak.cpm.k4$k4_temp10_1)/3), y= log2((peak.cpm.k4$k4_temp26_1 + peak.cpm.k4$k4_temp26_2 + peak.cpm.k4$k4_temp26_3)/3), pch = 20,xlab="H3K4me3 10ºC (CPM)",ylab="H3K4me3 26ºC (CPM)",main="26ºC vs 10ºC")
lines(x = c(0,20), y = c(0,20), type = "l", col="dodgerblue2"
      
```
```{r boxplot global cpm}

boxplot((peak.cpm.k4$k4_temp10_1 + peak.cpm.k4$k4_temp10_2 +peak.cpm.k4$k4_temp10_3)/3,
        (peak.cpm.k4$k4_temp14_1 + peak.cpm.k4$k4_temp14_2 +peak.cpm.k4$k4_temp14_3)/3,
        (peak.cpm.k4$k4_temp20_1 + peak.cpm.k4$k4_temp20_2 +peak.cpm.k4$k4_temp20_3)/3,
        (peak.cpm.k4$k4_temp26_1 + peak.cpm.k4$k4_temp26_2+peak.cpm.k4$k4_temp26_3)/3,
        outline=F,col=c("#00FFFF","#6699FF","#CC33FF","#FF0000"),ylab="CPM",names=c("10ºC","14ºC","20ºC","26ºC"),las=2,cex.lab=1)

wilcox.test(x = (peak.cpm.k4$k4_temp10_1 + peak.cpm.k4$k4_temp10_2 + peak.cpm.k4$k4_temp10_3)/3, y = (peak.cpm.k4$k4_temp14_1 + peak.cpm.k4$k4_temp14_2 +peak.cpm.k4$k4_temp14_3)/3,alternative = "less")

wilcox.test(x = (peak.cpm.k4$k4_temp10_1 + peak.cpm.k4$k4_temp10_2 +peak.cpm.k4$k4_temp10_3)/3, y = (peak.cpm.k4$k4_temp20_1 + peak.cpm.k4$k4_temp20_2 +peak.cpm.k4$k4_temp20_3)/3,alternative = "less")

wilcox.test(x = (peak.cpm.k4$k4_temp10_1 + peak.cpm.k4$k4_temp10_2 +peak.cpm.k4$k4_temp10_3)/3, y = (peak.cpm.k4$k4_temp26_1 + peak.cpm.k4$k4_temp26_2 +peak.cpm.k4$k4_temp26_3)/3,alternative = "less")

wilcox.test(x = (peak.cpm.k4$k4_temp14_1 + peak.cpm.k4$k4_temp14_2 +peak.cpm.k4$k4_temp14_3)/3, y = (peak.cpm.k4$k4_temp20_1 + peak.cpm.k4$k4_temp20_2 +peak.cpm.k4$k4_temp20_3)/3,alternative = "less")

wilcox.test(x = (peak.cpm.k4$k4_temp20_1 + peak.cpm.k4$k4_temp20_2 +peak.cpm.k4$k4_temp20_3)/3, y = (peak.cpm.k4$k4_temp26_1 + peak.cpm.k4$k4_temp26_2 +peak.cpm.k4$k4_temp26_3)/3,alternative = "less")



```


```{r}
library(NormalyzerDE)

dim(peak.cpm.k4)

write.table(x = peak.cpm.k4[4:15],file = "peak_cpm_k4.tsv" ,
            quote = F,row.names = F,
            sep = "\t")

design.k4 <- data.frame(sample=colnames(peak.cpm[4:15]),
                     group=c(rep("temp10",3),rep("temp14",3), rep("temp20",3),rep("temp26",3)))



write.table(x = design.k4,file = "normalyzer_design_k4_global.tsv",quote = F,row.names = F,
            sep = "\t")


normalyzer(jobName = "normalyzer_design_k4_global_new",designPath = "normalyzer_design_k4_global.tsv",
         dataPath = "peak_cpm_k4.tsv",outputDir = ".", omitLowAbundSamples = T)


normalized.peak.cpm.k4 <- read.table(file="normalyzer_design_k4_global_new/Quantile-normalized.txt", header=T)
head(normalized.peak.cpm.k4)


boxplot(normalized.peak.cpm.k4, outline=F, cex.lab=1.5)

peak.cpm.k4[,1]<- read.counts.k4.temp10c.1[,1]
peak.cpm.k4[,2]<- read.counts.k4.temp10c.1[,2]
peak.cpm.k4[,3]<- read.counts.k4.temp10c.1[,3]


dim(peak.cpm.k4)
dim(normalized.peak.cpm.k4)

normalized.peak.cpm.k4 <-  cbind(peak.cpm.k4[,1], peak.cpm.k4[,2], peak.cpm.k4[,3], normalized.peak.cpm.k4)
colnames(normalized.peak.cpm.k4)[1:3] <- c("chr", "start", "end")
head(normalized.peak.cpm.k4)



boxplot((normalized.peak.cpm.k4$k4_temp10_1 + normalized.peak.cpm.k4$k4_temp10_2 +normalized.peak.cpm.k4$k4_temp10_3)/3,
        (normalized.peak.cpm.k4$k4_temp14_1 + normalized.peak.cpm.k4$k4_temp14_2 +normalized.peak.cpm.k4$k4_temp14_3)/3,
       (normalized.peak.cpm.k4$k4_temp20_1 + normalized.peak.cpm.k4$k4_temp20_2 +normalized.peak.cpm.k4$k4_temp20_3)/3,
       (normalized.peak.cpm.k4$k4_temp26_1 + normalized.peak.cpm.k4$k4_temp26_2 +normalized.peak.cpm.k4$k4_temp26_3)/3,
      outline=F,col=c("#00FFFF","#6699FF","#CC33FF","#FF0000","blue"),ylab="CPM",names=c("10ºC","14ºC","20ºC","26ºC"),las=2,cex.lab=1)


```

```{r}

library(limma)

experimental.design.k4 <- model.matrix(~ -1+factor(c(1,1,1,2,2,2,3,3,3,4,4,4)))
colnames(experimental.design.k4) <- c("k4_temp10","k4_temp14","k4_temp20","k4_temp26")

experimental.design.k4

normalized.peak.cpm.k4 <- normalized.peak.cpm.k4[,4:15]

linear.fit.k4 <- lmFit(normalized.peak.cpm.k4, experimental.design.k4)

contrast.matrix.k4 <- makeContrasts(k4_temp14-k4_temp10,k4_temp20-k4_temp10,k4_temp26-k4_temp10,
                                 levels=c("k4_temp10","k4_temp14","k4_temp20","k4_temp26"))

contrast.linear.fit.k4 <- contrasts.fit(linear.fit.k4, contrast.matrix.k4)
contrast.results.k4 <- eBayes(contrast.linear.fit.k4)


```

```{r temp14c-temp10c contrast}
temp14c.vs.temp10c.k4 <- topTable(contrast.results.k4, number=nrow(normalized.peak.cpm.k4),coef=1,sort.by="logFC")
head(temp14c.vs.temp10c.k4)
peak.ids.14c.10c.k4 <- rownames(temp14c.vs.temp10c.k4) 
fold.change.14c.10c.k4 <- temp14c.vs.temp10c.k4$logFC
q.value.14c.10c.k4 <- temp14c.vs.temp10c.k4$adj.P.Val

activated.peaks.temp14c.vs.temp10c.k4 <- peak.ids.14c.10c.k4[fold.change.14c.10c.k4 > 1 & q.value.14c.10c.k4 < 0.05]
repressed.peaks.temp14c.vs.temp10c.k4 <- peak.ids.14c.10c.k4[fold.change.14c.10c.k4 < - 1 & q.value.14c.10c.k4 < 0.05]

length(activated.peaks.temp14c.vs.temp10c.k4)
length(repressed.peaks.temp14c.vs.temp10c.k4)

```

```{r temp20c-temp10c contrast}
temp20c.vs.temp10c.k4 <- topTable(contrast.results.k4, number=nrow(normalized.peak.cpm.k4),coef=2,sort.by="logFC")
head(temp20c.vs.temp10c.k4)
peak.ids.20c.10c.k4 <- rownames(temp20c.vs.temp10c.k4) 
fold.change.20c.10c.k4 <- temp20c.vs.temp10c.k4$logFC
q.value.20c.10c.k4 <- temp20c.vs.temp10c.k4$adj.P.Val

activated.peaks.temp20c.vs.temp10c.k4 <- peak.ids.20c.10c.k4[fold.change.20c.10c.k4 > 1 & q.value.20c.10c.k4 < 0.05]
repressed.peaks.temp20c.vs.temp10c.k4 <- peak.ids.20c.10c.k4[fold.change.20c.10c.k4 < - 1 & q.value.20c.10c.k4 < 0.05]

length(activated.peaks.temp20c.vs.temp10c.k4)
length(repressed.peaks.temp20c.vs.temp10c.k4)

```
```{r temp26c-temp10c contrast}
temp26c.vs.temp10c.k4 <- topTable(contrast.results.k4, number=nrow(normalized.peak.cpm.k4),coef=3,sort.by="logFC")
head(temp26c.vs.temp10c.k4)
peak.ids.26c.10c.k4 <- rownames(temp26c.vs.temp10c.k4) 
fold.change.26c.10c.k4 <- temp26c.vs.temp10c.k4$logFC
q.value.26c.10c.k4 <- temp26c.vs.temp10c.k4$adj.P.Val

activated.peaks.temp26c.vs.temp10c.k4 <- peak.ids.26c.10c.k4[fold.change.26c.10c.k4 > 1 & q.value.26c.10c.k4 < 0.05]
repressed.peaks.temp26c.vs.temp10c.k4 <- peak.ids.26c.10c.k4[fold.change.26c.10c.k4 < - 1 & q.value.26c.10c.k4 < 0.05]

length(activated.peaks.temp26c.vs.temp10c.k4)
length(repressed.peaks.temp26c.vs.temp10c.k4)

```

